<center> <h2>Interactive Python - Week 2</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 2.1 Event-Driven Programming*

**A. What is it?**

- not like a procedural programming
- not "linear" execution
    - e.g. not Run A --> B --> C etc.
- "any event can occur and the program will respond to it"
    - a function that runs when an event occurs is called a *handler*

**B. Types of Events**

- Input
    - Button
    - Text box
- Keyboard
    - key down
    - key up (release)
- Mouse
    - click
    - drag
- Timer

**C. Timer Events**

- have the timer fire periodically

```{r eval=FALSE, engine='python'}

import simplegui

## Event handler
def tick():
    print "tick!"

## Register handler
timer = simplegui.create_timer(1000, tick) ##run this ever thousand milliseconds

## Start it
timer.start()
    
```

- Make sure to make it stop!

**D. Event Queue**

- Nothing you can do to make two events not happen at the same time
- the system is queuing events as they occur, and executing code accordingly
- "you can't control this queue"
    - your program should be written to handle any order of stuff
    - you cannot handle multiple events at once
- if you make your handler really long, other events will sit and wait and the game will be unresponsive
- the "Hollywood model" of programming
    - "don't call us, we'll call you"

####*Video 2.2 Local v. global variables*

**A. The Deal**

- event handlers need to share information with each other
- handlers will share a common set of global variables

**B. The difference**

- when you define something outside of a function, it is a global variable
- when you define one inside a function, it is a local variable
- you need to be strategic about using global vs local
- you cannot modify a global variable inside a function

**C. Why use a local variable?**

- give a descriptive name to a quantity
- avoid computing something multiple times

**D. Generally speaking...**

- You should always prefer local variables unless you have a really specific reason to use global
- you can use the ```global``` command within a function to modify globals inside of functions

####*Video 2.3 Simplegui*

**A. What the hell is it?**

- this was built (in JavaScript) specifically for CodeSkulptor

**B Explanation**

```{r eval=FALSE, engine='python'}
# CodeSkulptor runs Python programs in your browser.
# Click the upper left button to run this simple demo.

# CodeSkulptor runs in Chrome 18+, Firefox 11+, and Safari 6+.
# Some features may work in other browsers, but do not expect
# full functionality.  It does NOT run in Internet Explorer.

import simplegui

message = "Welcome!"

# Handler for mouse click
def click():
    global message
    message = "Good job!"

# Handler to draw on canvas
def draw(canvas):
    canvas.draw_text(message, [50,112], 36, "Red")

# Create a frame and assign callbacks to event handlers
frame = simplegui.create_frame("Home", 300, 200)
frame.add_button("Click me", click)
frame.set_draw_handler(draw)

# Start the frame animation
frame.start()
```

- the frame is the window that pops up
- the canvas is where you can draw stuff

**C. Handlers**

- canvas has draw handlers
- frame has key, keyboard, other handlers
- you could bring in images and sounds

**D. Program Structure**

1. Define globals (state)
2. Define helper functions
3. Define classes
4. Define event handlers
5. create a frame
6. Register event handlers
7. start frame and timers
- example:

```{r eval=FALSE, engine='python'}
# Import the module
import simplegui

# Define global variables (program state)
counter = 0

# Define "helper" functions
def increment():
    global counter
    counter = counter + 1
    
def buttonpress():
    global counter
    counter = 0 ## button click will reset coutner to 0
    
# Create a frame
frame = simplegui.create_frame("SimpleGUI Test", 100, 100)

# Register event handlers
timer = simplegui.create_timer(1000, tick)
frame.add_button("Click Me!", buttonpress)

# Start frame and timers
frame.start()
timer.start()

```



