<center> <h2>Interactive Python - Week 7</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 7.1 - Acceleration and Friction*

**A. Notes**

- we will spend the last two weeks making Asteroids
- left key (rotate counterclockwise), right key (rotate clockwise)
    - speed is controlled by acceleration, not velocity
- adding friction:
    - ship generally slows if it isn't being accelerated
    
**B. Basic Physics**

- position update: ```position += velocity```
- velocity update: ```velocity += acceleration```
- variables for the ship class:
    - self.pos = ship's position (vector)
    - self.vel = ship's velocity (vector)
    - self.angle = ship's orientation (scalar)
    - self.thrust = is the ship accelerating? (boolean)
- velocity update - acceleration in direction of forward vector
    - forward = [math.cos(self.angle), math.sin(self.angle)]
    - if self.thrust:
    -   self.vel[0] += forward[0]
    -   self.vel[1] += forward[1]
- update position then velocity
- go back to this lecture to see more notes and sample code

####*Video 7.2 - Spaceship Class*

**A. Notes**

```{r eval=FALSE}
# Partial example code for Spaceship

import simplegui


class ImageInfo:
    def __init__(self, center, size, radius = 0, lifespan = None, animated = False):
        self.center = center
        self.size = size
        self.radius = radius
        if lifespan:
            self.lifespan = lifespan
        else:
            self.lifespan = float('inf')
        self.animated = animated

    def get_center(self):
        return self.center

    def get_size(self):
        return self.size

    def get_radius(self):
        return self.radius

    def get_lifespan(self):
        return self.lifespan

    def get_animated(self):
        return self.animated

    
# art assets created by Kim Lathrop, may be freely re-used in non-commercial projects, please credit Kim
    
# ship image
ship_info = ImageInfo([45, 45], [90, 90], 35)
ship_image = simplegui.load_image("http://commondatastorage.googleapis.com/codeskulptor-assets/lathrop/double_ship.png")

# sound assets purchased from sounddogs.com, please do not redistribute
ship_thrust_sound = simplegui.load_sound("http://commondatastorage.googleapis.com/codeskulptor-assets/sounddogs/thrust.mp3")


# Ship class
class Ship:
    def __init__(self, pos, vel, angle, image, info):
        self.pos = [pos[0],pos[1]]
        self.vel = [vel[0],vel[1]]
        self.thrust = False
        self.angle = angle
        self.angle_vel = 0
        self.image = image
        self.image_center = info.get_center()
        self.image_size = info.get_size()
        self.radius = info.get_radius()
        
    def draw(self,canvas):
        canvas.draw_circle(self.pos, self.radius, 1, "White", "White")

    def update(self):
        pass
```

- two images: with or without thrustster
    - you will also need to rotate based on velocity lecture

####*Video 7.3 - Sounds*

**A. Playing Sounds is Straightforward**

- the hardest part can be just finding your own sounds
- sounds are associated with codeskultptor, not the frame (so you need to reset!)
- implement in Chrome! (firefox sucks with sounds)

####*Video 7.4 - Sprites*

**A. Sprites**

- 2D visusal elements laid on top of the canvas to make things pretty
- now, sprites are "logical entities used to organize/represent images that add visual complexity to a game"
- sprite sheets consisted a collection of sprites organized as a single image. Note that the individual sprites need not be regularly spaced on the sprite sheet
- note: in this class, we load them as individual images

**B. Color and Transparency**

- RGB model - three red, green, blue channels
    - ```rgb(255,0,0)``` is "Red"
- we want to draw irregular shapes (like asteroids) that lie in rectangular images
    - the trick is to incorporate transparency
    - ```rgba(255,0,0,0.5)``` --> 1 is opaque, 0 is transparent
- create your images with transparent "alpha channel" in Photoshop, gIMP, paint.net, PNG image is popular
- GIMP - open-source photoshop
- we give the sprite custom draw and update methods...makes it easy to draw

####*Video 7.5 - Debuggin*

**A. Sound**

- not all browsers support the same file types
- no errors are raised if sound fails
- be sure you have non-zero volume!

**B. Code Style**

- put functions in dictionaries!

```{r eval=FALSE}
    # Avoiding long if/elif chain with dictionary mapping values to actions

def paddle1_faster():
    global paddle1_vel
    paddle1_vel += 2

def paddle1_slower():
    global paddle1_vel
    paddle1_vel -= 2
    
def paddle2_faster():
    global paddle2_vel
    paddle2_vel += 2    

def paddle2_slower():
    global paddle2_vel
    paddle2_vel -= 2


inputs = {"up": paddle2_slower,
          "down": paddle2_faster,
          "w": paddle1_slower,
          "s": paddle1_faster}

def keydown(key):
    for i in inputs:
        if key == simplegui.KEY_MAP[i]:
            inputs[i]()
```

- avoid repeated code!
    - use functions or methods, make it correct, and move on with your life

**C. Magic Constants**

- magic constant = very specific hard-coded values
    - it's much, much better to code in the calculations so your code reacts to changes and others can understand what you've done
- seriously bro...stop hard coding values
- give your constants names, as well
    - don't use x, j, y, etc.
    
**D. One More Thing**

- break your stuff down into many pieces, name those pieces so they're understandable

####*Video 7.6 - Spaceship*

**A. Spaceship**

- ship should rotate
- should thrust/accelerate in the forward direction
    - two images...thruster on, thruster off
    - thrust sound while accelerating
- missile velocity = combination of ship velocity and forward vector
- missile spawns from the tip of the ship
- asteroids spawn at random place
- 