<center> <h2>Interactive Python - Week 6</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 6.1 - Object-Oriented Programming (pt 1)*

**A. Notes**

- OOP = object containing data that has methods to do stuff
    - writing things an OOP way involves creating your own object types and methods
- Python convention is to capitalize classes 

```{r eval=FALSE}
class Character:
    def __init__(self, name, initial_health):
        self.name = name
        self.health = initial_health
        self.inventory = []
        
    def __str__(self):
        s  = "Name: " + self.name
        s += " Health: " + str(self.health)
        s += " Inventory: " + str(self.inventory)
        return s
    
    def grab(self, item):
        self.inventory.append(item)
        
    def get_health(self):
        return self.health
    
def example():
    me = Character("Bob", 20)
    print str(me)
    me.grab("pencil")
    me.grab("paper")
    print str(me)
    print "Health:", me.get_health()
    
example()

```

- ```__str__(self):``` should build up and then return a string with summary information
- defining your own types and methods can make big programs much simpler and more powerful

####*Video 6.2 - OOP (pt 2)*

**A. Continued**

```{r eval=FALSE, engine='python'}

# ball physics code for generic 2D domain
# the functions inside() and normal() encode the shape of the ennvironment

import simplegui
import random
import math

# Canvas size
width = 600
height = 400

# Ball traits
radius = 20
color = "White"

# math helper function
def dot(v, w):
    return v[0] * w[0] + v[1] * w[1]

class RectangularDomain:
    def __init__(self, width, height):
        self.width = width
        self.height = height
        self.border = 2

    # return if bounding circle is inside the domain    
    def inside(self, center, radius):
        in_width = ((radius + self.border) < center[0] < 
                    (self.width - self.border - radius))
        in_height = ((radius + self.border) < center[1] < 
                     (self.height - self.border - radius))
        return in_width and in_height

    # return a unit normal to the domain boundary point nearest center
    def normal(self, center):
        left_dist = center[0]
        right_dist = self.width - center[0]
        top_dist = center[1]
        bottom_dist = self.height - center[1]
        if left_dist < min(right_dist, top_dist, bottom_dist):
            return (1, 0)
        elif right_dist < min(left_dist, top_dist, bottom_dist):
            return (-1, 0)
        elif top_dist < min(bottom_dist, left_dist, right_dist):
            return (0, 1)
        else:
            return (0, -1)

    # return random location
    def random_pos(self, radius):
        x = random.randrange(radius, self.width - radius - self.border)
        y = random.randrange(radius, self.height - radius - self.border)
        return [x, y]

    # Draw boundary of domain
    def draw(self, canvas):
        canvas.draw_polygon([[0, 0], [self.width, 0], 
                             [self.width, self.height], [0, self.height]],
                             self.border*2, "Red")
        
class CircularDomain:
    def __init__(self, center, radius):
        self.center = center
        self.radius = radius
        self.border = 2
        
    # return if bounding circle is inside the domain    
    def inside(self, center, radius):
        dx = center[0] - self.center[0]
        dy = center[1] - self.center[1]
        dr = math.sqrt(dx ** 2 + dy ** 2)
        return dr < (self.radius - radius - self.border)

    # return a unit normal to the domain boundary point nearest center
    def normal(self, center):
        dx = center[0] - self.center[0]
        dy = center[1] - self.center[1]
        dr = math.sqrt(dx ** 2 + dy ** 2)
        return [dx / dr, dy / dr]
    
    # return random location
    def random_pos(self, radius):
        r = random.random() * (self.radius - radius - self.border)
        theta = random.random() * 2 * math.pi
        x = r * math.cos(theta) + self.center[0]
        y = r * math.sin(theta) + self.center[1]
        return [x, y]
        
    # Draw boundary of domain
    def draw(self, canvas):
        canvas.draw_circle(self.center, self.radius, self.border*2, "Red")
    
class Ball:
    def __init__(self, radius, color, domain):
        self.radius = radius
        self.color = color
        self.domain = domain
        
        self.pos = self.domain.random_pos(self.radius)
        self.vel = [random.random() + .1, random.random() + .1]
        
    # bounce
    def reflect(self):
        norm = self.domain.normal(self.pos)
        norm_length = dot(self.vel, norm)
        self.vel[0] = self.vel[0] - 2 * norm_length * norm[0]
        self.vel[1] = self.vel[1] - 2 * norm_length * norm[1]
    

    # update ball position
    def update(self):
        self.pos[0] += self.vel[0]
        self.pos[1] += self.vel[1]
        if not self.domain.inside(self.pos, self.radius):
            self.reflect()

    # draw
    def draw(self, canvas):
        canvas.draw_circle(self.pos, self.radius, 1, 
                           self.color, self.color)
        

# generic update code for ball physics
def draw(canvas):
    ball.update()
    field.draw(canvas)
    ball.draw(canvas)

field = RectangularDomain(width, height)
# field = CircularDomain([width/2, height/2], 180)
ball = Ball(radius, color, field)
        
frame = simplegui.create_frame("Ball physics", width, height)

frame.set_draw_handler(draw)

frame.start()

```

- OOP is much more flexible, enables you to quickly make substantive changes

####*Video 6.3 - Working With Objects*

**A. Using one class to create objects**

```{r eval=FALSE}

# Particle class example used to simulate diffusion of molecules

import simplegui
import random

# global constants
WIDTH = 600
HEIGHT = 400
PARTICLE_RADIUS = 5
COLOR_LIST = ["Red", "Green", "Blue", "White"]
DIRECTION_LIST = [[1,0], [0, 1], [-1, 0], [0, -1]]

# definition of Particle class
class Particle:
    
    # initializer for particles
    def __init__(self, position, color):
        self.position = position
        self.color = color
        
    # method that updates position of a particle    
    def move(self, offset):
        self.position[0] += offset[0]
        self.position[1] += offset[1]
        
    # draw method for particles
    def draw(self, canvas):
        canvas.draw_circle(self.position, PARTICLE_RADIUS, 1, self.color, self.color)
    
    # string method for particles
    def __str__(self):
        return "Particle with position = " + str(self.position) + " and color = " + self.color


# draw handler
def draw(canvas):
    for p in particle_list:
        p.move(random.choice(DIRECTION_LIST))
    
    for p in particle_list:
        p.draw(canvas)


# create frame and register draw handler
frame = simplegui.create_frame("Particle simulator", WIDTH, HEIGHT)
frame.set_draw_handler(draw)

# create a list of particles
particle_list = []
for i in range(100):
    p = Particle([WIDTH / 2, HEIGHT / 2], random.choice(COLOR_LIST))
    particle_list.append(p)

# start frame
frame.start()
```

- be sure to declae objects as your new method
- when we say ```print``` an object, Python looks for the ```__str__``` method of that class
- if you try to call a method that a class doesn't actually have, you'll get an AttributeError
- advice for efficient BlackJack: use the mini-project template
    - play it at the console first, then go to canvas
- functions vs methods:
    - functions are defined outside of classes
    - methods are defined inside of and part of classes

####*Video 6.4 - Classes for Blackjack*

**A. Rules**

- programming as an art, not a science
    - make some "design decisions"
- highest hand with 21 or less wins.
    - dealer wins ties
- dealer cards --> 1 down, 1 up
- players --> all up (?)

**B. Implementation**

- types you need:
    - card(rank, suit, image)
    - hand(collection of cards)
    - deck(collection of cards)
- a type is the data inside it and the behaviors of that type
- OO design --> primary objective is to facilitate code reuse
    - card class could be used for other games (like poker)
- abstraction --> separate how the object is implemented from how you use it
    - again, you want something that people can interface with for building their own stuff

**C. Overloading**

- overloading is a situation where you allow declarations of a class with different numbers of arguments
    - this is not supported in Python

####*Video 6.5 - Tiled images*

**A. Notes**

- the idea here is that if you have a bunch of regular-size icons you want (say cards), you can load a single image file and then subset out the card you want
- ```canvas.draw_image(tiled_image, card_pos, CARD_SIZE, ...)```
- order the cards in a specific image in this tiled image...we lean heavily on that order

**B. Code Example**

```{r eval=FALSE}

# demo for drawing using tiled images

import simplegui

# define globals for cards
RANKS = ('A', '2', '3', '4', '5', '6', '7', '8', '9', 'T', 'J', 'Q', 'K')
SUITS = ('C', 'S', 'H', 'D')

# card sprite - 950x392
CARD_CENTER = (36.5, 49)
CARD_SIZE = (73, 98)
card_image = simplegui.load_image("http://commondatastorage.googleapis.com/codeskulptor-assets/cards.jfitz.png")



# define card class
class Card:
    def __init__(self, suit, rank):
        self.rank = rank
        self.suit = suit

    def draw(self, canvas, loc):
        i = RANKS.index(self.rank)
        j = SUITS.index(self.suit)
        card_pos = [CARD_CENTER[0] + i * CARD_SIZE[0],
                    CARD_CENTER[1] + j * CARD_SIZE[1]]
        canvas.draw_image(card_image, card_pos, CARD_SIZE, loc, CARD_SIZE)

# define draw handler        
def draw(canvas):
    one_card.draw(canvas, (155, 90))

# define frame and register draw handler
frame = simplegui.create_frame("Card draw", 300, 200)
frame.set_draw_handler(draw)

# createa card
one_card = Card('S', '6')

frame.start()
```

**C. Using Dropbox**

- you're not required to do this...but it's fun
- just put the stuff on there and grab the "share" URL/link
- IMPORTANT: won't work with the normal url
    - you have to replace "www" with "dl"
    - e.g. "https:dl.dropbox.com/..."
- once the image is loaded once, it is cached
- to test this...make sure you run it on a different computer!
    - it might work on your computer just because you can download it while signed in to DropBox
    
####*Video 6.6 - Visualizing Objects*

**A. Notes**

```{r eval=FALSE}

##################
# Object creation and use
# Mutation with Aliasing

class Point1:
    def __init__(self, x, y):
        self.x = x
        self.y = y
    
    def set_x(self, newx):
        self.x = newx
    
    def get_x(self):
        return self.x

p = Point1(4, 5)
q = Point1(4, 5)
r = p

p.set_x(10)

print p.get_x()
print q.get_x()
print r.get_x()


##################
# Object shared state
# Mutation of shared state

class Point2:
    def __init__(self, coordinates):
        self.coords = coordinates
    
    def set_coord(self, index, value):
        self.coords[index] = value
    
    def get_coord(self, index):
        return self.coords[index]

coordinates = [4, 5]
p = Point2(coordinates)
q = Point2(coordinates)
r = Point2([4, 5])

# this line will change the "shared state" of coordinates, affecting p and 1
p.set_coord(0, 10)

print p.get_coord(0)
print q.get_coord(0)
print r.get_coord(0)


##################
# Objects not sharing state

class Point3:
    def __init__(self, coordinates):
        self.coords = list(coordinates)
    
    def set_coord(self, index, value):
        self.coords[index] = value
    
    def get_coord(self, index):
        return self.coords[index]

coordinates = [4, 5]
p = Point3(coordinates)
q = Point3(coordinates)
r = Point3([4, 5])

p.set_coord(0, 10)

print p.get_coord(0)
print q.get_coord(0)
print r.get_coord(0)
```

- ```__init__()``` method being called implicitly when we initialize an object

####*Video 6.7 - Programming Tips*

**A. Common Mistakes using Objects and Classes**

```{r eval=FALSE}
###################
# Broken code

class ball:
    def ball(pos, rad):
        position = pos
        radius = rad
        return ball
    
    def get_position():
        return position

b = ball([0,0], 10)

print get_position(b)



###################
# Fixed code

class Ball:
    def __init__(self, pos, rad):
        self.position = pos
        self.radius = rad
    
    def get_position(self):
        return self.position

b = Ball([0,0], 10)

print b.get_position()


##################
# Example while

def countdown(n):
    """Print the values from n to 0."""

    i = n
    while i >= 0:
        print i
        i -= 1

countdown(5)


##################
# Collatz

def collatz(n):
    """Prints the values in the Collatz sequence for n."""

    i = n
    while i > 1:
        print i
        
        if i % 2 == 0:
            i = i / 2
        else:
            i = 3 * i + 1
           
colnatz(1000)


#################
# Timeout

i = 1
while i > 0:
    i += 1
	
print "Done!"
```

- functions use the syntax ```function(object)```, methods take syntax ```object.method()```
- remember that all methods must have the argument "self"
- always include an ```__init__``` and probably a ```__str__```
    - ```__init__()``` does not need a ```return``` statement...it will implicitly just return an object of that class
- reference internal variables as ```self.variable_name``` when defining your class
- Python convention is to capitalize classes

**B. Better Looping**

- For looping is a bit rigid
- a sometimes-better construction is the ```while``` loop
- don't write infinite loops, yo

####*Video 6.8 - Blackjack*

**A. Notes**

- values of hands:
    - numbers = number
    - face card = 10
    - Ace = 1 or 11
- no aces --> easy
- multiple aces --> you never count more than one ace as 11!
- dealing with Aces
    - when computing the value of the player's hand, take the max of (value with ace as 11, value with ace as 1), assuming ace as 11 not bust

```{r eval=FALSE}
If the hand has no Aces
    return hand_value
else
    if hand_value + 10 <= 21
        return hand_value + 10
    else
        return hand_value
```

- for the dealer, you'll need to use a while loop
- something like:

```{r eval=FALSE}
while dealer_hand.get_value() < 17:
    dealer_hand.hit()
```

- in our version, dealer wins ties
- keep things very simple here...follow the project description
- List of prime numbers in a range:

```{r eval=FALSE}
## the "Sieve of Eatosthenes"
# Computes primes less than n

n = 1000
numbers = range(2,n)

results = []
while len(numbers) != 0:
    
    # Add the first number in numbers to the end of results
    results.append(numbers[0])
    print results
    
    # Remove every number in numbers that is evenly divisible by...
    #	the number that you ahd just added to results
    to_remove = [value for value in numbers if (value % numbers[0])==0]
    for value in to_remove:
        numbers.remove(value)
    
print results
print len(results)
```