<center> <h2>Exploratory Data Analysis - Week 2</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 2.1 Lattice Plotting System (pt 1)*

**A. Why Use Lattice?**

- useful for lots of plots at once
- good for high dimensional data

**B. The Lattice Plotting System**

- lattice contains graphics functions like ```xyplot()```, ```bwplot()```, and ```levelplot()```
- you create the entire plot in a single function call

**C. Main Functions**

- ```xyplot``` - main function for scatterplots
- ```bwplot()``` - box-and-whiskers plots (boxplots)
- ```histogram()``` - histograms
- ```stripplot()``` - like a boxplot but with actual points
- ```dotplot()``` - plot dots on "violin strings"
- ```splom()``` - scatterplot matrix (like ```pairs()``` in base plotting system)
- ```levelplot()```, ```contourplot()``` for "image" data

**D. Simple Lattice Plot**

```{r sec1lattice, eval=FALSE}
library(datasets)
library(lattice)
airquality <- transform(airquality, Month = factor(Month))
xyplot(Ozone ~ Wind | Month, data = airquality, layout = c(5,1))
```

**E. Lattice Behavior**

- base graphics functions return an object of class trellis, which is then printed
    - this is typically automatic though
- you can save this in an object (and call it when you need it)
    - it would be better to save the code and data (plots take lots of memory)

####*Video 2.2 Lattice Plotting System (pt 2)*

**A. Lattice Panel Functions**

- lattice has panel functions which control what happens inside each panel of thep lot
- you can make you own panel functions if you want too

**B. A custom panel function call**

```{r sec2lattice, eval=FALSE}
## Custom panel function
xyplot(y ~ x | f, panel = function(x, y, ...){
    panel.xyplot(x, y, ...) ## first call default panel function
    panel.lmline(x, y, col = 2) ## Overlay a simple linear regression line
})
```

- you can't use annotation functions from base system
    - in fact, you can pretty much never mix across plotting systems
- you can look at a TON of data

**C. Summary**

- latticep lots are constructed with a single function call
- margin, labels, spacing all automatically handled
- this is ideal for creating conditioning plots where you examine the same kind of plot under many conditions
- panel functions can be specified/customized to modify what is plotted in each of the plot panels

####*Video 2.3 ggplot2 (pt 1)*

**A. What is ggplot2?**

- a package in R based on Grammar of graphics
- a "third" graphics system in R
- you have a grammar (verbs, nouns, adjectives) of graphics
- the idea is to build plots by combining different pieces
- better documentation at [ggplot2 website](http://ggplot2.org)
- "theory" of graphics on which to build new graphics and graphics objects
- "shorten the distance from mind to page"

**B. Grammar of Graphics**

- mapping, aesthetic attributes (color, shape, size), geometric objects (points, lines, bars)

**C. The Basics: qplot()**

- works much like the ```plot()``` function in base graphics system
- looks for data in a data frame, similar to lattice, or in the parent environment
- plots are made up of aesthetics (size, shape, color) and geoms (points, lines)
- factors are important for indicating subsets of the data
    - they should be labeled
    - the labels will be plotted
- ```ggplot()``` is the core function and very flexible for doing things qplot() cannot do

####*Video 2.4 ggplot2 (pt 2)*

**A. Modifying Aesthetics**

- you can change color in a plot by another factor
    - e.g. ```qplot(disply, hwy, data = mpg, color = drv)```

**B. Adding a geom**

- a smoother ('loess')
    - ```qplot(disply, hwy, data = mpg, geom = c("point", "smooth"))
    
**C. Histograms**

- just specify a single variable
- you can use the "fill=" argument to show a histogram but fill up the histogram

**D. Facets**

- similar to panels in lattice
- use facets = {factor name} to set up how you want the panels to display

**E. Density smooth**

- ```qplot(log(eno), data = maacs, geom="density"
- two separate distributions (density plots):
    - ```qplot(log(eno), data=maacs, geom="density", color = mopos)```
- some scatterplots:
    - again, you can use color= to get thep lots to change color
    - you could also use shape = {factor} to give the points different shapes by group

**F. You could split out with Facets**

- ```qplot(log(pm25), log(eno), data = maacs, geom = c("point", smooth"), method = "lm", facets = .~ mopos)```

**G. Summary of qplot**

- the ```qplot()``` function has lots of built-in features
- syntax somewhere between base and lattice
- produces very nice graphics, essentially publication ready
- difficult to customize the plot aesthetically from ```qplot()```

####*Video 2.5 ggplot2 (pt 3)*

**A. Basic COmponents of a ggplot2 Plot**

- a data frame
- aesthetic mapping (how data are mapped to color, size)
- geoms - geometric objects like points, lines, shapes
- facets - for conditional plots
- stats - statistical transformations like binning, quantiles, smoothing
    - smoothing, fitting, confidence intervals
- scales - what scale an aesthetic map uses
- coordinate system

**B. Building plots with ggplot2 (no qplot)**

- add stuff piece by piece (layer by layer)
    - plot the data
    - overlay a summary
    - metadata and annotation
- you can store a ggplot2 plot in an object

####*Video 2.6 ggplot2 (pt 4)*

**A. Adding More Layers: Smoothers**

- get a trend
- let's get an OLS trendline:
- ```g + geom_point() + geom_smooth(method = "lm")``` 
- facets are used to make conditioning plots
    - note that ggplot is going to take the labels from the factors

**B. Annotation**

- labels: ```xlab, ylab, labs, ggitle```
- each of the geom functions has options to modify
    - you can use the theme() function to change lots of stuff
- two standard appearance themes are included: ```theme_gray()```, ```theme_bw()```

**C. Modifying Aesthetics**

- ```geom_point(color = "steelblue", size = 4, alpha = 1/2)```
- when you use ```aes()``` to assign color to a variable, you're doing something fundamentally different than using a constant like "black"

**D. Modifying labels**

- use the ```labs()``` function for modifying titles and x-, y-axis labels

**E. Customizing the Smoother**

- you could make the lines thicker, you could add a line type
    - inside ```geom_smooth()```, use se= FALSE to take out confidence intervals

**F. Changing the Theme**

- use the theme functions to change plot and stuff like that

####*Video 2.7 ggplot2 (pt 5)*

**A. Notes About Axis Limits**

- in the base plot system, you could set ylim to truncate outliers
- in ggplot, it's going to plot all the data (even with outliers)
    - changing the y limits with ylim in ggplot will actually subset the data (removing outliers)
    - you need to use ```coord_cartesian(ylim = c(,)``` to get base plot-esque truncation

**B. Conditioning on a Continuous Variable**

- the ```cut()``` function returns a factor variable breaking things into the quantiles you want

**C. Summary**

- ggplot2 is very powerful and flexible if you learn the grammar and the various elements
- many more types of plots can be made; explore and mess around with the package