<center> <h2>Programming for Everybody - Week 6</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 6.1 - Strings*

**A. Introduction to Strings**

- you can look inside strings like you would lists
    - characters are index starting at 0
- you can also do slicing
- the ```len()``` function returns the number of characters (including spaces)
- we can loop through strings
    - bits of strings can be accessed with the index

```{r eval=F, engine='python'}
fruit = "banana"
for letter in fruit:
    print letter
```

**B. Looping**

- find largest letter, smallest letter
- count certain letters
- you can concatenate with ```+```

**C. String Methods***

- ```str1.lower()``` returns a lower-case COPY of the string
- ```str1.upper()``` returns an upper-case COPY of the string (doesn't modify the string)
- you can use ```dir(<objectname>)``` to return all the methods built into certain types
- the ```find()``` function returns the index of a substring within another string
    - note that it only finds the first instance, and that indexing starts at 0
    - returns a "-1" if it doesn't find it
    - find takes an optional second argument telling it where to start looking
- the ```replace()``` function is like a search and replace operation

```{r eval=F, engine='python'}
greet = 'Hello Bob'
nstr = greet.replace('Bob','Jane') ##returns a COPY with replacement
```

- ```lstrip()``` and ```rstrip()``` strip whitespace from the left and right, respectively
- ```strip()``` removes all the outer whitespace
- you can search prefixes (case matters!) using the ```startswith``` METHOD
- Python is good at "ripping data into pieces"
- searching a bunch of stuff in text is called "parsing"

####*Video 6.2 - Bonus: Monash Museum of Computing History*

**A. Notes**

- located in Sydney, Australia
- it's a danger with computing students...that they will focus too much only on "the latest"
- the first computers were people (and, more specifically, women)
- history can be quickly forgotten
    - we may end up spending effort solving sub-problems whose solutions were embedded in past technologies


