<center> <h2>Sabermetrics 101 - Week 1</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Reading 1.1 - Dictionary Definitions*

**A. Dictionary Definitions**

- Bill James coined the term Sabermetrics in 1980
- sabermetrics = "the application of statistical analysis to baseball records, especially in order to evaluate and compare the performance of individual players"

####*Video 1.1 - What is Sabermetrics? Pt 1*

**A. Definitions**

- evaluating players, evaluating teams
- "sabermetrics often employs more complex statistical categories than those used in traditional baseball statistics"
- [Google Ngram Viewer](https://books.google.com/ngrams):
    - sabermetrics used more frequently in the last decade
- word first used in 1980

####*Reading 1.2 - Bill James's Refined Definition*

**A. Notes**

- "good sabermetrics respects the validity of all types of evidence"
- be skeptical, be non-commital
- Sabermetrics is about "why teams win, how the game changes, and why the game changes, and why the game works"

####*Video 1.2 - What is Sabermetrics? Pt 2*

**A. More Definitions**

- sabermetrics by its nature is unemotional, non-committal
- good sabermetrics by its nature is unemotional, non-commital
- sabermetrics is "the science of baseball"
- sabermetrics = study of the game of baseball through observation and experimentation (when applicable)

####*Video 1.3 - What is Statistics? (pt 1)*

**A. Notes**

- Bill James says "I have never been a statistician"
- "statistics" as a field of scholarship; the science of collecting, classigyin, summarizing, and understanding data using math and data analytics
- Henry Chadwick as an important early sabermetrician
- two types of stats:
    - Counting Stats (e.g. total runs)
    - Rate states (e.g. runs/game)

####*Video 1.4 - What is Statistics? (pt 2)*

**A. Notes**

- measures of central tendency:
    - mean, median, mode
- measures of variability/dispersion:
    - standard error of the mean, standard deviation of the dataset
- exploratory data analysis (EDA):
    - histograms, boxplots
    - scatterplots (bivariate analysis)
    - correlation
- always start with a scatterplot
- regression to the mean:
    - need to be aware of this in studying baseball
    
####*Video 1.5 - what is Data Science?*

**A. Why even care?**

- data science is sexy
    - the field is expanding rapidly
- this course is like an "introduction to applied data science"
    - the application is baseball
- what is it:
    - more than data analysis
    - requires an understanding of databases
    - an understanding of client needs
    - an understanding of privacy and ethics
- BU SQL Sandbox will be a key tool in this
- "the Drew Conway Data Science Venn Diagram"
- science is essentially having an intense curiosity about the natural world and observing it carefully
- reminder: this course is an applied introduction to data science

####*Reading 1.3 - Introduction to Big Data*

**A. Notes**

- "the amount of data generated by people these days is staggering"
- a few bits of trvia:
    - In August 2012, Facebook reported it was processing 500+ TB of new data per day
    - In October 2012, the internet archive reached 10 PB (PB = 1000 TB = 1000000 GB)
    - In February 2014, IBM said daily data created exceeded 2.3 exabytes (EB = 1000 PB = 1000000 TB = 1000000000 GB)
- data analytics = using data for discovery and analysis
- "big data" refers to datasets whose size is beyond the ability of typical database software tools to capture, store, manage, and analyze
    - might also refer to data which move too fast for normal processing tools to keep up

####*Video 1.6 - What is Big Data?*

**A. Notes**

- MLB recently announced [Statcast](http://m.mlb.com/news/article/119234412/statcast-primer-baseball-will-never-be-the-same)
- some data comes into databases easier, because the technology generating the data was designed and engineered with data collection in mind
- lots of data exist in unstructured, potentially undocumented formats
- big data don't always lead to better analytics
    - you might not get better answers just from throwing more data at the problem
- careful observation is important to data science
- big datasets still have missing data
- you will invariably need to merge incompatible datasets
    - "think before you merge"

####*Reading 1.4 - Introduction to Databases*

**A. Notes**

- database = data (values) and schema (structure)
- database management systems (DBMS) = define new schema, insert/modify/delete data, retrieve data
- in this course, we use the [Lahman Database of Baseball Statistics](http://www.seanlahman.com/baseball-archive/statistics/)

**B. Why not just use Excel?**

- limitations:
    - record size (Excel only allows ~1 million rows, and gets slow
    - Excel is only built for 1 user
    - types of information (Excel is designed for numeric data)
    - report generation (custom reports get complicated fast)

**C. Why Databases, then?**

- larger record sizes
- many users/editors
- different data types (numbers, text, dates can live together)
- flexible report generation (databases are designed to easily run complex queries)

####*Video 1.7 - What is a Database?*

**A. Intro to Databases**

- SQL is a specific programming language for relational databases (RDMBS)
- relational DBs are more flexible than navigational DBs
    - set of tables and relationships
    - SQL is a RDBMS
- asking a question = a SQL query
- teams want analysts who know SQL
- we'll be using Oracle MySQL
- three kinds of databases:
    - hierarchical
    - network (graph)
    - relational
- lots and lots of different SQLs
    - SQL is open source
- MySQL isn't great for big data or unstructured data
    - there are better DBMS for unstructured data

####*Video 1.8 - The Lahman Database*

**A. What is it?**

- put on the internet in the 1990s by Sean Lahman
- [Lahman documentation](http://seanlahman.com/files/database/readme2014.txt)
- this is a relational database (there are a bunch of tables)
- Master table:
    - unique player ID
    - player background (age, birth country)
- the unique keys help us to join tables
- each record has a unique index within the tables
- the description of tables and their contents and relationships = "schema" or "database model"
    - the relational table setup saves storage space (you don't end up storing lots of 0s, like in an Excel sheet)
    
####*Video 1.9 - SQL Sandbox Walkthrough*

**A. Intro to SQL**

- you can run queries right inside the sandbox (Lahman Database exploration mode)
- the command ```SHOW TABLES``` will list all the tables in the database
    - this will also give you how many rows were returned by the query
- to learn about a certain table, use ```DESCRIBE <tablename>```
    - returns a list of the fields with variable type, primary keys, etc.
    - you can download whatever you query to an Excel sheet

####*Reading 1.5 - Query Grading*

**A. Ungraded vs. Graded Queries**

- there is an exploration mode of the sandbox, but it will also have graded instances after videos
    - try running in exploration mode before submitting for credit
- partial credit is given on queries

**B. Using the Data**

- we have read-only privileges for the data in the Sandbox (no table creation)
- however, you could download a local copy [here](http://courses.edx.org/c4x/BUx/SABR101x/asset/lahman-sql_2013-12-10.zip)

####*Video 1.10 - First SQL Sandbox Query*

**A. Notes**

- get all the fields in the batting table, where the year is 1998:

```{r eval=FALSE}
SELECT *
FROM batting
WHERE yearID = 1998
```

- the "download results" button gives you a CSV with the result of your query
- SQL is not case-sensitive

**B. W3schools documentation**

- [SQL syntax](http://www.w3schools.com/sql/sql_syntax.asp)
- important commands:
    - ```SELECT``` extracts data from a database
    - ```UPDATE``` updates data in a database
    - ```DELETE``` deletes data from a database
    - ```INSERT INTO``` inserts new data into a database
    - ```CREATE DATABASE``` creates a new database
    - ```ALTER DATABASE``` modifies a database
    - ```CREATE TABLE``` creates a new table
    - ```ALTER TABLE``` modifies a table
    - ```DROP TABLE``` deletes a table
    - ```CREATE INDEX``` creates an index (search key)
    - ```DROP INDEX``` deletes an index
- operators for the WHERE clause:
    - ```=``` Equal
    - ```<>``` Not equal. In some version of SQL, this may be ```!=```
    - ```>,<``` greater than, less than
    - ```>=, <=```, greater than or equal, less than or equal
    - ```BETWEEN``` between an inclusive range
    - ```LIKE``` search for a pattern
    - ```IN``` speecify multiple possible values for a column
    
####*Reading 1.6 - Exploring the Lahman Database Schema*

**A. Notes**

- docementation for the database available [here](http://seanlahman.com/files/database/readme2012.txt)
- [list of MySQL data types](http://www.w3schools.com/sql/sql_datatypes.asp)

####*Video 1.11 - More Queries*

**A. Continuing**

- SQL expects character variables in single quotes
    - e.g. ```WHERE teamID = 'SEA' AND yearID = 1998;```
- you can use AND and OR and NOT in the "where" clause to zoom in on just what you want
- you can combine AND and OR clauses, using parentheses to form complex expressions
    - see [here](http://www.w3schools.com/sql/sql_and_or.asp) for more
    
####*Video 1.12 - Even More Queries*

**A. Runs per Game**

- Chadwick reported and talked about this a lot
- you can select specific stuff using a comma-delimited list in the SELECT statement
    - you can even use formulas in this call!
- check it out:
    
```{r eval = FALSE}
SELECT playerID, yearID, teamID, G, R,
    R/G AS RPG
FROM Batting
WHERE teamID = 'NYA' AND yearID = 2013;
```

- the ability to do these calculations in-line is powerful

####*Video 1.13 - Chadwick (pt 1)*

**A. Henry Chadwick**

- the earliest sabermetrician
    - known as the "Father of Base Ball"
- a writer and scorekeeping pioneer
- early in his career, Chadwick wrote about cricket for the New York newspapers
- elected to the baseball hall of fame in 1938
- inventor of the box score, author of the first rule book in 1858
- the first box score ever published was in 1845
- original things tracked: "hands out" and runs
- he changed the path of the game through his writing, and rule-making, and promotion efforts
- a walk used to be considered a "base on error" (an error by the pitcher)

####*Video 1.14 - Henry Chadwick (pt 2)*

**A. More Notes on the Scoresheet**

- he wanted to recerate the events of the game so others could read them
- the long-form scoresheet could be turned into a publishable box score for the papers
- this is where sabermetrics becomes science
    - Chadwick helps to record and organize "observations"
- without Henry Chadwick, the game of baseball today would be very different

####*Video 1.15 - Interview with David Cameron*

**A. Notes**

- David Cameron, from FanGraphs
- sabermetrics as an art?
    - all sciences have art
    - this is about ideas and the fostering of ideas
    - you will always have personal interpretations of data
- the true test of validity will always be predictive accuracy
- you learn the most from your own mistakes
    - want to be a write? keep writing
    - build a portfolio to show people

####*Video 1.16 - Interview with Jonah Keri*

**A. Notes**

- Johan Keri (lead baseball write, Grantland.com)
- MoneyBall was not the only driver of analytics in MLB front offices
- initial slow uptake might have been a self-preservation instinct
- smart business practices took over:
    - this opened the door for stats and analytics to complement scouting
- people are really excited about MLBAM
- big name they keep mentioning: Rob Neyer
- 