<center> <h2>Principles of Computing - Week 6</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 6.1 - Introduction*

**A. Notes**

- we hope you can "use a computer to do interesting and exciting things"

####*Video 6.2 - Search*

**A. Search**

- spatial search:
    - 1st-person shooter have computer-controlled AI
    - Google Maps does a spatial search when providing routes
- search:
    - simpler: key-value lookup
    - complex: complex space of paramters, alignments
    
####*Video 6.3 - Generators*

**A. Notes**

- programmatically generating a sequence of values to iterate over them
    - they don't produce the entire sequence at once...just produce one value at a time, explicitly intended for iteration


**B. Examples**

- a generator for the max function:

```{r eval=FALSE}
print "max in gen:", max(num*2-3 for num in range(7))
```

- what does it mean for a function to ```yield``` something?
    - the function returns a value, but remembers its state
    - it goes back to this state each time you call it
- diff between ```yield``` and ```return```?
    - return says you aren't coming back
    - yield says you'll be coming back to continue

####*Video 6.4 - Stacks and Queues*

**A. Notes**

- two important data structures in compsci
- used to hold data during program execution

**B. Stacks**

- LIFO structure --> Last In, First Out
- push element = put it in
- pop = take something out

**C. Queues**

- FIFO structure --> first in, first out
- put something in --> enqueue
- take something out --> dequeue

**D. An Example**

- "do I want the most recent thing I pushed, or do I want things to come back out in the order I pushed them in?"
- when you call ```len(x)```, Python is calling that object's ```__len__``` method
- ```__iter__``` yields stuff, acts like a generator
- using a queue or stack class rather than a list helps you to control the ways that data are added/pulled out of the object

####*Video 6.5 - Inheritance*

**A. Notes**

- a mechanism by which one class can inherit the methods of another class
- a "sub-class" inherits from a "base class"
- this use of sub-classes helps you to get around writing big, ugly if/elif blocks
- queue's and stacks are "restricted access containers" (RAC)
- Python does not pay attention to types very well

**B. Duck-Typing**

- How does Python handle figuring out class?
    - "if it looks like a duck and quacks like a duck, it's a duck"

**C. Simple Inheritance**

- code example:

```{r eval=FALSE}
# Simple inheritance

class Base:
    def hello(self):
        print "hello"
        
    def message(self, msg):
        print msg
        
class Sub(Base):
    def message(self, msg):
        print "sub:", msg
        
obj = Sub()
obj.hello()
obj.message("what's going to happen?")

baseobj = Base()
baseobj.hello()
baseobj.message("another message")
```

- redefining a method in the subclass overrides the method from the base class

```{r eval=FALSE}
"""
Simple example of using inheritance.
"""

class Base:
    """
    Simple base class.
    """    
    def __init__(self, num):
        self._number = num

    def __str__(self):
        """
        Return human readable string.
        """
        return str(self._number)
        
class Sub(Base):
    """
    Simple sub class.
    """
    def __init__(self, num):
        Base.__init__(self,num)
    
obj = Sub(42)
print obj
```

- Python doesn't automatically call the initializer of the base class

####*Video 6.6 - Grid Class*

**A. Notes**

- a grid is a partition of a rectangular space into cells
- grids enable you to model a geometric region as a 2D data structure
- indexing is in the form of (row,col)
- ```four_neighbours`` = "rook continguity"
- ```eight_neighbours``` = queen contiguity
- an example: Conway's Game of Life
    - "a cellular automaton"
    - state of the grid changes every iteration

**B. Game of Life Example**

- some code:

```{r eval=FALSE}

"""
Game of Life demo
"""

import poc_grid
import poc_gol_gui

# constants
EMPTY = 0 
FULL = 1

class GameOfLife(poc_grid.Grid):
    """
    Extend Grid class to support Game of Life
    """
    
    def update_gol(self):
        """
        Function that performs one step of the Game of Life
        """
        
        updated_grid = [[self.update_cell(row, col) \
                            for col in range(self.get_grid_width())] \
                            for row in range(self.get_grid_height())]
        
        for col in range(self.get_grid_width()):
            for row in range(self.get_grid_height()):
                if updated_grid[row][col] == EMPTY:
                    self.set_empty(row, col)
                else:
                    self.set_full(row, col)
                    

    def update_cell(self, row, col):
        """
        Function that computes the update for one cell in the Game of Life
        """
        # compute number of living neighbors
        neighbors = self.eight_neighbors(row, col)
        living_neighbors = 0
        for neighbor in neighbors:
            if not self.is_empty(neighbor[0], neighbor[1]):
                living_neighbors += 1
            
        # logic for Game of life        
        if (living_neighbors == 3) or (living_neighbors == 2 and not self.is_empty(row, col)):
            return FULL
        else:
            return EMPTY
        
# run gui        
poc_gol_gui.run_gui(GameOfLife(30, 40))
```

####*Video 6.7 - Grid Search*

**A. Notes**

- let's think about a grid as modeling a geometric domain
- simple strategy for finding something on a 1d path:
    - just traverse the path, and eventually you'll pass what you're looking for

**B. 2D search**

- model the way a fire burns
- this is breadth-first search
- the boundary is a queue

**C. Breadth-First Search**

- search ends when the queue is empty
- this can be generalized to a graph (a bunch of nodes connected by edges)

####*Video 6.8 - Zombie Apocalypse*

**A. Notes**

- humans flee zombies, zombies stalk humans
- overview --> [HERE](https://class.coursera.org/principlescomputing2-004/wiki/zombie)
- keep track of where are the humans, where are the zombies
- phase one:
    - implement the methods of the zombie class
- distance field:
    - find all of the closest cells and how many steps away they are
    - main task of this mini-project
- zombie class inherits from the Grid class
- the main challenge is implementing breadth-first search