<center> <h2>Principles of Computing - Week 7</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 7.1 - The Importance of recursion*

**A. Notes**

- a Russian nesting doll is a good example of recursion:
    - just follow one rule
    - the problems get smaller as you go
- "recursion marks the transition from being a programmer to being a computer scientist"
- binary search is an example of recursion:
    - repeat the same process over and over until you find what you want
    - NOT strictly equal to looping, because you are iterating toward an answer systematically and monotonically

####*Video 7.2 - Recursion*

**A. Notes**

- you define a function that calls itself
- each time you call it, it has less data than you had originally
- best utilized when a problem exhibits common sub-structures
    - problem is a composition of similar, smaller subproblems
- steps:
    - identify the base case
    - identify the recursive case

**B. An Example**

```{r eval=FALSE}
def sum_up_to_one(n):
    base_case: 1 when n = 1
    recursive case: n + sumupto(n-1) when n > 1
```

- recursive = inductive
- recursion example:

```{r eval=FALSE}
# Recursion

def sum_up_to(n):
    if n == 1:
        # base case
        return 1
    else:
        # recursive case
        return n + sum_up_to(n - 1)

print sum_up_to(1)
print sum_up_to(2)
print sum_up_to(10)
print sum_up_to(55)

print sum(range(56))

def is_palindrome(word):
    if len(word) < 2:
        # base case
        return True
    else:
        # recursive case
        if word[0] != word[-1]:
            return False
        else:
            return is_palindrome(word[1:-1])

print is_palindrome("a")
print is_palindrome("aa")
print is_palindrome("is")
print is_palindrome("madamimadam")
print is_palindrome("abcdefdcba")
```

- recursion problems require careful decomposition of the problem into base and recursive cases

####*Video 7.3 - Binary Search*

**A. Notes**

- the simplest "divide an conquer" algorithm

**B. Example**

- you have some SORTED list L of numbers and want to check if number K is in it
- naive approach...check every element for equality
    - "linear search"
    - double list size, double running time
- binary search run time $log_{2}(n)$
    - much slower growth than linear time
    - only works on a sorted list
- recursion is always the same process...you change the bits of data it's applied to
    - and those bits keep getting smaller and smaller
    
####*Video 7.4 - Visualizing Recursion*

**A. Notes**

- Viz mode executes and builds up a trace of execution
- recursive function builds up a "call stack"
    - stack of functions which depend on each other's output
    - "a sequence of recursive calls waiting to be executed"
    - need a LIFO method
    
**B. QuickSort**

- an n_log_n sorting algorithm

####*Video 7.5 - Recurrences*

**A. Notes**

- a recurrence is a simplified recursive function where we already have the values of some base cases
- a recurrence is a set of equations
- solution:
    - closed form function returning roughly the same values as the recurrence
- you can always just look up the closed form solutions to common recurrences

####*Video 7.6 - Reading Files*

**A. Notes**

- CodeSkulptor runs on a web page, so it can't read files off your harddrive
- we'll be reading files over the network

```{r eval=FALSE}
# Reading files from the network

import urllib2
import codeskulptor

FILENAME = "examples_files_dracula.txt"

url = codeskulptor.file2url(FILENAME)
netfile = urllib2.urlopen(url)

#data = netfile.read()
#print data
#print type(data)

#for line in netfile.readlines():
#    print "line: ", line[:-1]
```

- on your desktop, you'd call ```myfile = open(FILENAME, "r")```

####*Video 7.7 - Word Wrangler*

**A. Notes**

- the only way to get better at recursion is to actually use it
- we're going to do ordered lists WITHOUT built-in support
- the goal is to build a game that creates a list of words that can be made out of the letters in an input word

**B. Merge Sort**

- a divide an conquer algorithm
- break big list into smaller ones
- sort those smaller ones

**C. Gen_All_Strings**

- think recursively:
    1. split the input word into "first" and "rest"
    2. generate all strings out of rest
    3. add all strings that include first at any position in the generated strings
- when you already know your input is ordered or sorted, you can write more powerful programs
- recursion can help you solve complex problems in an elegant way