<center> <h2>Intro to Matlab</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video M.1 - Matlab variables*

**A. Notes**

- variables created with the assignment operator, ```=```
    - must have only numbers, letters, underscores
    - cannot start with a number

```{r eval=FALSE, comment='%'}
%create a variable called x
x = 4

% Get the avlue pi
g = pi^2
```

- default behavior is to store results of calcs in ```ans```

####*Video M.2 - Matlab as a Calculator*

**A. Notes**

- "sooner or later, we'll encounter a problem whose solution requires a complicated formula and some serious number crunching"
- operators:
    - addition ```+```
    - subtraction ```-```
    - grouping ```()```
    - square root ```sqrt()```
- remember order of ops

####*Video M.3 - Creating Vectors*

**A. Notes**

- storing a vector in Matlab

```{r eval=FALSE}
%row vector
x = [-2,-1,0,1,2]

% col vector
y = [-2;-1;0;1;2]

% line plot
plot(x,y)

% Regularly spaced vectors
% x = startVal:spacing:end
x = -2:0.1:2;

%regularly spaced col vector
x = (-2:2)';' 
```

- When using the ```:``` operator, the vector will include all values in the range "up to but not including" the last number
- using the colon, spacing of 1 is assumed unless specified
the ```'``` is the transpose operator in Matlab

####*Video M.4 - Calculations with Vectors*

**A. Notes**

- Matlab = MATtix LABoratory
- the operators are designed to work with matrices
- elementwise operations are performed with ```.*```, ```./``` and ```.^```
- scalars are automatically expanded to match the size of the vector

####*Video M.5 - Matlab Functions*

**A. Notes**

- there are thousands of Matlab functions
- the standard syntax is ```y = function(x)```
- matlab functions are more general than mathematical functions
- the ```min()``` function returns not only the minimum but also the index of that minimum
- functions return multiple outputs by asking for them in ````[]```

####*Video M.6 - Matrices*

**A. Notes**

- how to do this:

```{r eval=FALSE}
%define a 2 by 4 matrix
m = [1,2,3,4; 5, 6,7,8]

% Three ways to get a 2 by 3 matrix
% The first way: 
A1 = [1,4;2,5;3,6;]

% The second way:
x = [1;2;3];
y = [4;5;6];
A2 = [x,y]

% The third way:
z = 1:6;
A3 = reshape(z, [3,2]) 
```

- in Matlab, the comma is like ```cbind()``` in R and the semicolon is like ```rbind()```

####*Video M.7 - Matrix Multiplication*

**A. Notes**

- matrix multiplication is fundamental to the software controlling devices with sensors
- for matrices to be multipliable...internal dimensions have to be the same
    - dimensions of the result will be outer dimensions
- in Matlab, we use the ```*``` operator
- order matters!

####*Video M.8 - Accessing Elements of a Vector*

**A. Notes**

- Matlab indexing starts at 1
- get the fourth element of v with ```a = v(4)```
- vectors/matrices are mutable...you can use indexing to change values or even to add values:
    - adding a 6th element to a vector of length 5 would b ```v(6) = 7```

####*Video M.9 - Accessing Elements of an Array*

**A. Notes**

- subsetting is by ```a1 = M(rownum, colnum)```
- you can use logical vectors for subsetting too (or really any vector of indexes)
- could even use ```a1 = M(1:4, 2)``` for the first four rows of column 2
- to refer to a complete column, use ```col2 = M(:,2)```

####*Video M.9 - String Variables*

**A. Notes**

- working with strings

```{r eval=FALSE}
item = 'oil' %1x3 array of type char.expand(
country = 'CM'

%characters stored as string variables
file = [item, country, '.mat'];

% above creates file name, below reads it
LOAD(file)
```

####*Video M.10 - Logical Operators*

**A. Notes**

- Matlab uses the ```&``` operator
    - not equal ```-=```
    - equal to ```==```
    - and ```&```
    - or ```|```
    - not ```-```
    
####*Video M.11 - Conditional Data Selection*

**A. Notes**

- you can use a logical vector to subset stuff in Matlab

```{r eval=FALSE}
indx = v < 5
v(indx) = NA
```

####*Video M.12 - Reshaping Arrays*

**A. Notes**

- the ```reshape()``` functions takes three args:
    - vectors to be reshaped
    - number of rows
    - number of cols
- output arrays are filled column by column

####*Video M.13 - Multiple Plots*

**A. Notes**

- create a lineplot using ```plot(month, avgTemp, 'r:o')```
- the default behavior of "plot" is to overwrite the most recent plot with the new one

```{r eval=FALSE}
%multiple type plots
bar(month, maxTemp, 'yellow') %1. Plot bar graphs
hold on %2. Plot the next thing over it
plot(month, avgTemp, 'red') %overlay the line plot
```

####*Video M.14 - Making Histograms*

**A. Notes**

- pass a vector to the ```HIST``` function
    - the function will also give you the output variables counts and centers if you want
- manually define 20 bins --> ```HIST(my_vector, 20)```
- you can also specify SPECIFIC bins:
    - ```HIST(my_vector, <some vector of numbers>)```

####*Video M.15 - Stem Plots*

**A. Notes**

- great for showing "shape of the signal as well as its sample rate"
- accepts two inputs:
    - time values of samples (x)
    - signal values (y) 
- this is accomplished with the function ```STEM(x,y)```

####*Video M.16 - The For Loop*

**A. Notes**

- a for loop "allows repeated execution of commands for a fixed number of times
- MATLAB syntax:

```{r eval=FALSE}
y(1) = 1
for n = 1:6
    y(n+1) = y(n) - 0.1*y(n)
end
```

- remember that you can control the step in-line, as in ```for 1:2:21``` to count from 1 to 21 by twos

####*Video M.17 - If Else Statements*

**A. Notes**

- great for piecewise functions
- syntax

```{r eval=FALSE}
if x<0
    f = 1 - x
else
    f = 2 + x
end
```