<center> <h2>Using Python to Access Web Data - Week 5</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 5.1 - Web Services Overview*

**A. Notes**

- HTML really not written to provide applications data
- XML and JSON are more made for this purpose
    - data expected to be used by applications
- Python Dictionary --> Java HashMap
    - "the wire protocol"
- Python dictionary --> wire format --> Java HashMap
- \textbf{serialization} - creating wire format from an internal structure
- \textbf{de-serialization} - creating an internal structure in some language from the wire format
- XML came soon after HTML
    - very structured, self-describing data
- JSON came later

####*Video 5.2 - Understanding the REST Architecture*

**A. Notes**

- REST = "how the web should work"
- the web as a network of standards

####*Video 5.3 - eXtensible Markup Language*

**A. Notes**

- representing arrays is better in JSON
- documents probably better in XML
    - MS Office, as example
- a tree of nodes, "elements"
- a complex element is a node containing other nodes
- tags can have attributes
- self closing tag has no values...just start tag with an attribute, ending in a slash
- indentation is irrelevant
- whitespace is generally discarded
- you can use an XML "pretty printer" to get better indentation
- similar to HTML syntax, but you create arbitrary tags
- \textbf{serialization} - converting data in one program into a common format that can be stored and/or transmitted between systems in a programming language-independent manner
- XML is a tree-like data structure
- each node can have three things as children:
    - other nodes
    - attributes
    - textual values

####*Video 5.4 - XML Schema*

**A. Notes**

- schema = "proper shape"
- what is legal XML?
- XML validation --> checking if the validation matches the schema
- a schema = "formalizing a relationship"
- W3C = "World Wide Web Consortium"
    - their common XML schema is called "XSD"
- an example of what we're talking about below
- XML schema:

```{r eval=FALSE}
<xs:element name="person">
    <xs:complexType>
        <xs:sequence>
            <xs:element name="full_name" type="xs:string" minOccurs="1" maxOccurs="1" />
            <xs:element name="child_name" type="xs:string" minOccurs="0" maxOccurs="10" />
        </xs:sequence>
    </xs:complexType>
</xs:element>
```

- and some valid XML:

```{r eval=FALSE}
<person>
    <full_name>Tove Refsnes</full_name>
    <child_name>Hege</child_name>
    <child_name>Stale</child_name>
    <child_name>HJim</child_name>
    <child_name>Borge</child_name>
</person>
```

**B. A note on Dates**

- the computer kind of prefers the ISO 8601 date/time format:
    - yyyy-mm-ddThh:mm:ssZ
- where Z gives the time zone
- natural (easy) sort if all the dates have exactly the same number of columns
- times/dates usually stored internally at UTC and then converted just at the moment where we show them to people

####*Video 5.5 - Parsing XML in Python*

- we will use ```xml.etree.ElementTree as ET``` to get a good XML parser for Python
- parsing = "deserialization"
    - "I don't know how I got this, btu I want to make sense of it
- creates a "tree" object with various methods for extracting data
- you could find all instances of some nested structure with something like ```mytree.findall('parent/child')```
    - returns a set of XML nodes (with all the same methods available)

####*Video 5.6 - BONUS: Ian Horrocks - RDG/OWL (advanced)*

**A. Notes**

- OIL = ontology interchange logic
- DAML = DARPA Agent Markup Language
- OWL = Web Ontology Language
- RDF = "resource description framework"
- 