---
title: "Assault Incidence in Seattle Not Impacted by Weather"
author: James Lamb
output: html_document
---

## Introduction

For this short report, I sought to explore the relationship between weather and assaults using crime data recorded in the city of Seattle, WA in the Summer of 2014.

The crime data come from the city of Seattle's [open data portal](https://data.seattle.gov/). The specific subset of the data used in this assignment was provided by the professor and is available [here](https://raw.githubusercontent.com/uwescience/datasci_course_materials/master/assignment6/seattle_incidents_summer_2014.csv). Weather data used below were pulled from [Weather Underground](http://www.wunderground.com/) using the [weatherData](https://cran.r-project.org/web/packages/weatherData/index.html) package in R.

**Note**- The raw R Markdown behind this post can be found at [this GitHub repo](https://github.com/jameslamb/MOOC_Stuff/blob/master/Communicating_Data_Science_Results/Assignments/UWCOMM_Assignment1.rmd).

```{r loadpackages, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE,cache=TRUE}
# Set up the environment
pkgs <- c("readr", "data.table", "ggplot2", "googleVis", "rCharts", "weatherData", "reshape2", "plyr", "rChartsCalendar")
lapply(pkgs, require, character.only = TRUE)

# Clean up workspace
rm(pkgs)
```


```{r getdata, eval=TRUE, echo=FALSE, warning=FALSE,message=FALSE, include=FALSE, cache=TRUE}

# Seattle Crime Data
SEA_data <- read_csv('https://raw.githubusercontent.com/uwescience/datasci_course_materials/master/assignment6/seattle_incidents_summer_2014.csv')
SEA_data <- as.data.frame(SEA_data)

# Seattle Weather Data
weather_data <- weatherData::getWeatherForDate("SEA", start_date="2014-06-01", end_date="2014-08-31", 
                            opt_detailed=TRUE,opt_all_columns=TRUE)

```

```{r preproc, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE,include=FALSE,cache=TRUE}

# Keep only the assault records
indx <- grep(pattern=".*asslt.*", x=SEA_data[,"Offense Type"],
             ignore.case=TRUE)
SEA_data <- SEA_data[indx,]


# Date Columns as Dates
date_cols <- list("Date Reported", "Occurred Date or Date Range Start","Occurred Date Range End")

for (colname in date_cols) {
    
    # zero-out the minutes and seconds in all the date cols (matching weather on hour is good enough)
    tmp <- data.frame(SEA_data[,colname])
    tmp <- sapply(X=tmp, FUN=function(x){gsub(":..:", ":00:", x)},
                  USE.NAMES=FALSE)
    
    print(colname)
    #COnvert to POSIXct with Pacifc Time
    tmp <- as.POSIXct(tmp, format="%m/%d/%Y %I:%M:%S %p",
                        tz="America/Los_Angeles")
    
    #Convcert to UTC
    tmp <- format(tmp, tz="UTC")
    SEA_data[,colname] <- as.POSIXct(tmp, format="%Y-%m-%d %H:%M:%S", tz="UTC")
    
    print("converted")
}

# Just keep date and hour
tmp <- sapply(X=weather_data$Date, FUN=function(x){gsub(":..:", ":00:", x)},
                  USE.NAMES=FALSE)
weather_data$DateUTC <- as.POSIXct(tmp, format = "%Y-%m-%d %H:%M:%S"
                ,tz="UTC")


# Add weather data to the Seattle data
SEA_data$Conditions <- NA;
SEA_data$Humidity <- NA;
SEA_data$TemperatureF <- NA;
SEA_data$Wind_SpeedMPH <- NA;

# For this, merge on date of occurence
for (i in c(1:nrow(SEA_data))) {
    tmpdate <- SEA_data[i,"Occurred Date or Date Range Start"]
    weather <- weather_data[(weather_data$DateUTC == tmpdate),]
    weather <- weather[nrow(weather),]
    
    for (colname in c("Conditions", "Humidity", "TemperatureF", "Wind_SpeedMPH")) {
        SEA_data[i,colname] <- as.character(weather[1,colname])
    }
}

# Assign a new factor for aggravated vs. non-aggravated
get_agg <- function(x){
    ifelse((regexpr("-AGG-",x) > 0), "Aggravated", "Non-aggravated")
}
SEA_data$Assault_Type <- sapply(X=SEA_data[,"Offense Type"], FUN="get_agg");

# Set the appropriate column classes
SEA_data$Conditions <- as.factor(SEA_data$Conditions)
SEA_data$Humidity <- as.numeric(SEA_data$Humidity)
SEA_data$TemperatureF <- as.numeric(SEA_data$TemperatureF);
SEA_data$Wind_SpeedMPH <- as.numeric(SEA_data$Wind_SpeedMPH);
SEA_data$Assault_Type <- as.factor(SEA_data$Assault_Type)

# Rename the key columns
names(SEA_data) <- gsub(" ", "_", names(SEA_data))
names(SEA_data) <- gsub("/", "_", names(SEA_data))
names(SEA_data) <- gsub("Occurred_Date_or_Date_Range_Start", "Date",
                        names(SEA_data))

```


## Exploratory Analysis

More specifically, I set out to create some simple visualizations showing the relationship between key weather variables and the incidence of assaults.

Figure 1 below shows the frequency and composition of assaults by weather conditions in the hour that the assault took place. It's obvious from this chart that non-aggravated assaults are far more likely to occur than aggravated assaults (thankfully!), but beyond that there is little evidence that weather conditions contribute strongly to the prevalence of assaults. Notice that "Clear" was the second most prevalent condition at the time of assaults. In a dataset this messy, I'm not comfortable saying that the difference between the frequency of assaults in "Clear" and "Mostly Cloudy" conditions is meaningful.

```{r stacked_bar, eval=TRUE, echo=FALSE, warning=FALSE, cache=TRUE,message=FALSE, fig.align='center'}

# Plot 1: GGplot stacked bar by conditions
melt_df <- reshape2::melt(SEA_data, id.vars = "Assault_Type",
                          measure.vars = "Conditions")
melt_df$Conditions <- melt_df$value
melt_df$value <- 1
summary_df <- aggregate(melt_df$value, by = list(melt_df$Assault_Type, melt_df$Conditions), FUN = "sum")
names(summary_df) <- c("Assault_Type", "Conditions", "Freq")
summary_df$Conditions <- with(summary_df, reorder(Conditions, Freq, FUN="sum"))

g1 <- ggplot(summary_df, aes(x=Conditions, y=Freq,fill=Assault_Type)) + 
    geom_bar(stat="identity") +
    ggtitle("Figure 1. Assault Frequencies by Weather Condition\n") +
    theme(plot.title = element_text(face="bold", size = 18),
          axis.title = element_text(size=13)) +
    coord_flip()
g1
```

Next, I sought to examine the relationship between daily temperature and assault frequency. There is some evidence in the criminology literature (see [here](http://www.ncbi.nlm.nih.gov/pubmed/16060743), for example) that hotter temperatures should correspond with higher incidences of violent crime.

In the figure below, I stack a time-series line chart of daily assault counts and a bar chart of daily average temperature. Based on my previous knowledge of criminology and anecdotal evidence, I'd expected to see at least a moderate relationship between temperature and crime. 

The (admittedly messy) data shown in these figures do not provide support for such a relationship.

```{r simple_ts, eval=TRUE, echo=FALSE, warning=FALSE, cache=TRUE,message=FALSE, fig.align='center', results='asis'}

# Compose two charts:
#   1. Daily line chart of assault frequencies
#   2. Daily average temperatures

# Create a day variable
SEA_data$MonthDay <- as.POSIXct(format(SEA_data$Date, format="%Y-%m-%d"), tz="UTC")

# Aggregate Assaults by date
SEA_data$ones <- 1
assault_df <- stats::aggregate(x=SEA_data$ones, by=list(SEA_data$MonthDay, SEA_data$Assault_Type), FUN = "sum")
names(assault_df) <- c("Date", "Assault_Type", "Freq")
cast_df <- reshape2::dcast(assault_df, Date ~ Assault_Type)

l1 <- gvisLineChart(cast_df, "Date", c("Aggravated", "Non-aggravated"), 
                       options = list(
                           title = "Fig 2a. Assault Frequency",
                           legend = "bottom",
                           width = 800,
                           height = 475
                       ))

# Get average temperature by date
temp_df <- aggregate(x=SEA_data$TemperatureF, by = list(SEA_data$MonthDay), FUN = "mean")
names(temp_df) <- c("Date", "Avg_Temp")
c1 <- gvisColumnChart(temp_df, xvar = "Date", yvar = "Avg_Temp",
                     options = list(
                         title = "Fig 2b. Daily Average Temperature (F)",
                         legend="none",
                         series = "[{color: 'orange'}]",
                         width = 800,
                         height = 300
))

# Merge the two charts
g1 <- gvisMerge(l1, c1, horizontal=FALSE)
print(g1,'chart')
```

## Conclusion

This short analysis suggests no strong relationship between weather conditions and assault incidence in the 2014 Seattle crime data. Much more work would need to be strengthen the claim, but this initial inquiry suggests that researchers hoping to model crime should start with non-weather factors.
