<center> <h2>Developing Data Products - Weeks 3 & 4</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 3.1 R Packages*

**A. What is an R Package?**

- a mechanism for e xntending the basic functionality of R
- a collection of R functions or other (data) objects
- organized in a systematic fashion to provide a minimal amount of consistency
- written by users/developers everywhere

**B. Where are These R Packages**

- CRAN and Bioconductor
- Available fro mGitHub, Bitbucket, Gitorious
- You don't have to put a package on a central repository, but doing so makes it easier for others to install your package

**C. What is the Point?**

- Documentation/vignettes
    - sharing forces you to document
- interface definition / clear API
- users know that it will at least load properly
- the upfront focus on structure makes it easier to maintain

**D. Package Development Process**

- write some code in an R script
- want to make code available to others
- incorporate R script file into R package structure
- write documentation for user functions
- include some other material (examples, demos, datasets, tutorials)
- hosting your package gives you a bunch more eyes checking that everything is working

####*Video 3.2 R Packages (pt 2)*

**A. R Package Essentials**

- An R package is started by creating a directory with the name of the R package
- A DESCRIPTION file which has info about the package
- R code in the R/ sub-directory
- documentation in the man/ sub-directory
- NAMESPACE

**B. The DESCRIPTION File**

- Package:
- Title:
- Description: 
- Version:
- Author:
- Maintainer:
- License:
- Depends:
- Suggests:
- Date:
- URL:
- Other fields could be added

**C. R Code**

- Copy R code into the R/ sub-directory
- There can be any number of files in this directory
- Usually separate out files into logical groups

**D. The NAMESPACE File**

- list the dependencies
- indicate which functions are exported (public to the user)
- non-exported functions can't be called directly by the user
    - make things cleaner by hiding implementation details
- you can also import functions from other packages

**E. The NAMESPACE File**

- Key directives
    - export()
    - import()
    - importFrom()
- you can also create new classes and methods that could be used on those classes

**F. Documentation**

- Documentation files (.Rd) place in man/ sub-directory
- written in a specific markup language
- required for every exported function
- you can document other things as well

**G. Building and Checking**

- R CMD build is a command-line program that creates a package archive file (.tar.gz)
- R CMD check runs a battery of tests on the package
- you can run R CMD build or R CMD check from the command-line using a terminal or command-shell application

**H. Checks you have to pass to be put on CRAN**

- R CMD check runs a battery of tests
- Documentation exists
- code can be loaded, no major coding problems or errors
- run examples on documentation
- check docs match code
- All tests must pass to put package on CRAN

**I. Getting Started**

- the ```package.skeleton()``` function in the utils package creates a skeleton R package
    - the directory structure will all be created
- If there are functions visible in your workspace, it writes R code files to the R/ directory

**J. Summary**

- R packages provide a systematic way to make R code available to others
- Standards ensure that packages are correct and usable
- the process:
    - create a new directory with R/ and man/ sub-directories
    - Write a DESCRIPTION file
    - Copy R code into the R/ sub-directory
    - Write a NAMESPACE file with exports/imports
    - build and check
    
####*Video 3.3 Build R Packages Demo (in RStudio)*

**A. What We're Doing**

- make a simple package with a function for a prediction algorithm that predicts high-dimensionality data using only the 10 best predictors

**B. Getting Started**

- New Project --> New Directory --> R Package
- you can write the documentation directly within the R code file (RStudio will strip out what is documentation and what is functional code)

**C. BUilding it**

- Click to the "Build" tab on the right in RStudio (only viewable if building a package)
- All documentation is created using ```roxygen```
- check out the packagename.Rd file to see the documentation you created earlier
- you can click the "Check" button on the right (under "BUild") to run R CMD Check

####*Video 3.4 R Classes and Methods*

**A. Basic Idea**

- you write new classes and methods to deal with new data types that R doesn't handle well by default

**B. Classes and Methods**

- A system for doing object-oriented programming
- R was originally quite interesting because it is both interactive and has a system for object orientation
    - other OOP languages (C++, Java, Lisp, Python, Perl) generally not interactive
- The code for submitting classes/methods allows the following:
    - idea that people start off as users and then become programmers/developers

**C. Two styles of classes and methods**

- S3 classes/methods
    - Included with version 3 of the S language
    - not rigorous or very formal
    - older
- S4 classes/methods
    - more formal and rigorous
    - also called new-style classes/methods
    - around since R 1.4.0

**D. Two worlds living side by side**

- both classes will live together for the forseeable future
- each can be used fairly independently of the other
- developers should use S4
- S3 are tempting...quick and dirty
- The code for implementing S4 classes/methods in R is in the methods package, which is usually loaded by default

**E. Object Oriented Programming in R**

- class = "a description of a thing"
- object = "an instance of a class"
- method = "a function that only operates on a certain class of objects"
- a generic function is an R function which dispatches methods (plot, mean, predict...)
- generic functions don't technically 'do anything'...just dispatch methods

**F. Stuff to Look Up**

- The help files for the ```methods``` package are extensive...read them

**G. Generics/Methods in R**

- S4 and S3 style generic functions look different but conceptually, they are the same
- you can write new methods for existing generics or new methods for new generics

**H. Methods**

- generic functions just call methods
- try this to see all the methods for a function:

```{r methods}
mean

## See S3 methods
methods("mean")

## See S4 Methods
showMethods("show")
```

**I. Generic/Method Mechanism**

- the first argument of a generic function is an object of a particular class
    1. The generic function checks to the class
    2. A search is done to see if there is an appropriate method
    3. If there exists a method for that class, then that method is called on the object
    4. If not, looks for a default and tries that
    5. IF no default, it breaks

**J. Examining Code for Methods**

- You cannot just print the code for a method like other functions because the code for the method is usually hidden
- if you want to see S3, use getS3method
- for S4, use ```getMethod()```

####*Video 3.5 R Classes and Methods (pt 2)*

**A. Calling Methods Directly**

- you could call the default directly
    - don't do this
- you get cleaner code by just calling the function and letting it choose the appropriate method

**B. WRite your own methods**

- If you write new methods for new classes, you'll probably end up writing methods for the following:
    - print/show
    - summary
    - plot

**C. S4 Classes**

- Why would you want to create a new class?
    - represent new types of data
    - new concepts/ideas that haven't been thought of yet
    - Abstract/hide implementation details from the user. Say things are "new" as in R does not know about them (not new to the statistical community)
    
**D. S4 Class/Method: Creating a New Class**

- A new class can be defined using the ```setClass``` function
    - at a minimum you need to specify the name of the class
    - you can also specify data elements called "slots"

**E. Summary**

- developing classes and associated methods is a powerful way to extend the functionality of R
- Classes define new data types
- methods extend generic functions to specify the behavior of generic functions on new classes
- as new data types and concepts are created, classes/methods provide a way for you to develop an intuitive interface for users

####*Videos 3.6 and 3.7 Deploying Prediction Algorithms to the Web with yhat*

**A. What is this**

- take you function to create a back-end algorithm
- host it somewhere and let other people build a front end
    - web developers can tap into your api (which you've created using yhat)

**B. Opening the Dataset**

- you can use the ```dir()``` to view all files/folders in current directory

**C. Set up the model**

- write some stuff for reading, cleaning, subsetting the data

**D. The Prediction Script**

- Just write some functions

**E. Uploading To yhat**

- Three functions to write
- see below:

```{r yhatr_example, eval=FALSE}
## Load packages
library(yhatr)

## load dependencies (library loadings )
model.require <- function(){
    library(fields)
}

## Take input data frame as an argument
## used to transform data before feeding to prediction model
model.transform <- function(df){
    df
}

## do the prediction
model.predict <- function(df){
    pollutant(df)
}

## deploy to yhat
yhat.config <- c(username = "jaylamb20@gmail.com",
                 apikey = "<api_key>",
                 env = "http://sandbox.yhathq.com")

## deploy your code to yhat
yhat.deploy("my_stuff")
```

**F. Reminder**

- yhat lets your stuff be accessed via an API
- you can also build a website or javascript (a web fron-end around this)

**G. Running a Yhat Model on the Web**

- the input code in "Documentation" has to be in JSON (a simple but ugly interface fory our stuff)

**H. Running a yhat Model from R**

- you can access the api from within R
    - let people call functions from within R without having to share all the data
    - people interface with your yhat stuff and just get what they need
- use ```yhat.predict``` to predict on your data

```{r yhatpredict, eval=FALSE}
## make sure you've configured!
library(yhatr)
yhat.config <- c(username = "jaylamb20@gmail.com",
                 apikey = "<api_key>",
                 env = "http://sandbox.yhathq.com")


## Send your dataframe "df", which just contains RHS variables, to the yhat algorithm "pollutant"
yhat.predict("pollutant",df)

## this will return a dataframe with the following:

    ## all your right hand side variables
    ## the model results (predictions)
```

- this is a really nice, lightweight way to make a prediction function available to people without sharing a huge dataset with them

**I. "Running a yhat model with cURL**

- this is something you can embed in a perl shell

**J. Summary**

- that was three simple ways to access a yhat algorithm
- There are a bunch of other ways (for example, through a web front end)
    - you could build a web front end on top of an R-based back end
    

