<center> <h2>Developing Data Products - Week 1</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 1.1 Introduction to Data Products*

**A. What is a Data Product?**

- an app, software, reproducible report based on data
- an R package would be a reproducible report too

**B. What We'll learn**

- How to use RStudio presenter
- a shiny app (web app)
- rCharts and interactive graphics

####*Video 1.2 Motivating Shiny*

**A. Three Problems**

- your software is hard to use
    - create an R package for installation?
- you want numbers updated automatically every month?
    - create a slidify presentation that pulls in new numbers
- web app for a prediction function?
    - make a Shiny app (even though you have minimal web app skills)

####*Video 1.3 Shiny 1 Introduction to Shiny*

**A. What is Shiny?**

- a platform for creating interactive R programs in a web page
- put up an algorithm where people can pick and click to use it
- use Shiny to create simple, yet powerful, web-based interactive data products in R
    - minimal dev cost
- Shiny is made by R Studio

**B. Some Mild Prerequisites**

- a little knowledge of html, css and js is very helpful
    - know html at the bare minimum
    - html gives web page structure and sectioning
    - css gives the style
    - js (JavaScript) gives you interactivity
- there are tons of tutorials online for these things
- Shiny uses bootstrap style (it looks nice, even on mobile)

**C. What Else is out there?**

- creating any solution requiring fairly deep knowledge of web client/server programming
- OpenCPU is a really neat project providing an API for calling R from web documents
    - he even hosts an OpenCPU server, but you can create your own
- you need a lot more knowledge of web programming for this

**D. Context**

- you created some stuff, let users on a website put in data and see what comes out
- you want to create a web site so that users can input the relevant predictors and obtain their prediction

**E. Getting Started**

- install R
- need Rtools if you're on Windows
- install shiny
- great tutorial at http://rstudio.github.io/shiny/tutorial/
- also get familiar with rstudio manipulate

**F. A Shiny Project**

- A shiny project is a directory containing at least two parts:
    - one named ui.R (for user interface) controls how it looks
    - one named server.R that controls what it does
- these files need to be in the same directory

**G. A minimal ui.R function**

```{r uimin, eval=FALSE}
library(shiny)
shinyUI(pageWithSidebar(
    headerPanel("Data Science FTW!"),
    sidebarPanel(
        h3('Sidebar text')  ## third level html heading
    ),
    mainPanel(
        h3('Main Panel text')
    )
))
```

**H. server.r example**

```{r servmin, eval=FALSE}
library(shiny)
shinyServer(
    function(input,output) {
        }
    )
```

**I. To run it**

- in R. change to the directories with these files and type runApp()
- or put the path to the directory as an argument
- it should open a browser window with the app running

####*Video 1.4 Shiny 2 basic html and getting input*

**A. R function for HTML Markup**

- the below gives some examples for HTML output (this goes in ui.R)

```{r htmltest, eval=FALSE}
shinyUI(pageWithSidebar(
  headerPanel("Illustrating markup"),
  sidebarPanel(
      h1('Sidebar panel'),
      h1('H1 text'),
      h2('H2 Text'),
      h3('H3 Text'),
      h4('H4 Text')

  ),
  mainPanel(
      h3('Main Panel text'),
      code('some code'),
      p('some ordinary text') ## normal text
  )
))

```

**B. Illustrating inputs ui.R**

```{r morehtml, eval=FALSE}
shinyUI(pageWithSidebar(
  headerPanel("Illustrating inputs"),
  sidebarPanel(
    numericInput('id1', 'Numeric input, labeled id1', 0, min = 0, max = 10, step = 1), ## id1 is the html label, printed label is next, default value, min, max, step value
    checkboxGroupInput("id2", "Checkbox", ## a 3-value check box
                   c("Value 1" = "1",
                     "Value 2" = "2",
                     "Value 3" = "3")),
    dateInput("date", "Date:")  
  ),
  mainPanel(  ## leave this blank

  )
))
```

**C. Another Part of ui.R - Grabbing Output**

```{r grabout, eval=FALSE}
 mainPanel(
        h3('Illustrating outputs'),
        h4('You entered'),
        verbatimTextOutput("oid1"),
        h4('You entered'),
        verbatimTextOutput("oid2"),
        h4('You entered'),
        verbatimTextOutput("odate")
  )
```

- where do those oid variables come from? the server function

```{r oidserv, eval=FALSE}
shinyServer(
  function(input, output) {
    output$oid1 <- renderPrint({input$id1})
    output$oid2 <- renderPrint({input$id2})
    output$odate <- renderPrint({input$date})
  }
)
```

####*Video 1.5 Shiny 3 Creating a very basic prediction function*

**A. Let's build our predict function**

```{r predictfunc, eval=FALSE}
shinyUI(
  pageWithSidebar(
    # Application title
    headerPanel("Diabetes prediction"),

    sidebarPanel(
      numericInput('glucose', 'Glucose mg/dl', 90, min = 50, max = 200, step = 5),
      submitButton('Submit') ## Adds a submit button. So there won't be auto-updating (only updates when you hit this)
    ),
    mainPanel(
        h3('Results of prediction'),
        h4('You entered'),
        verbatimTextOutput("inputValue"),
        h4('Which resulted in a prediction of '),
        verbatimTextOutput("prediction")
    )
  )
)

## put the prediction algorithm in the server:

diabetesRisk <- function(glucose) glucose / 200

shinyServer(
  function(input, output) {
    output$inputValue <- renderPrint({input$glucose}) ## print the input value
    output$prediction <- renderPrint({diabetesRisk(input$glucose)}) ## print the result of running this function
  }
)
```

####*Video 1.6 Shiny 4 Working with images*

**A. Image example**

- let's build an example with an image
- how to create a histogram of data
    - put a lisder on so that the user has to guess the mean

**B. ui.R**

```{r histexam, eval=FALSE}
shinyUI(pageWithSidebar(
    headerPanel("Example plot"),
    sidebarPanel(
        sliderInput('mu', 'Guess at the mean',value = 70, min = 62, max = 74, step = 0.05,)  ## give people a slider
    ),
    mainPanel(
        plotOutput('newHist') ## defined in server. newHist is the label for the output created in server.R
    )
))
```

**C. server.R**

```{r histexample, eval=FALSE}
library(UsingR)
data(galton)

shinyServer(
    function(input, output) {
        output$newHist <- renderPlot({
            hist(galton$child, xlab='child height', col='lightblue',main='Histogram') ## CREATE A HISTOGRAM
            mu <- input$mu  ## GRAB THAT INPUT GUESS
            lines(c(mu, mu), c(0, 200),col="red",lwd=5) ## DRAW LINES AT INPUT VALUE
            mse <- mean((galton$child - mu)^2)  ## TELL THE PERSON THE MSE
            text(63, 150, paste("mu = ", mu))  ## SHOW WHAT THE GUESS WAS
            text(63, 140, paste("MSE = ", round(mse, 2)))  ## SHOW WHAT MSE WAS
        })
        
    }
)
```

- how does this all work?
    - ui tells shiny to create a slider
    - slider creates input value
    - input value passed to server.R
    - server.R gives it the instructions back to render the plot (and how to mark it up)

####*Video 1.7 Shiny 5 Discussion*

**A. Tigether control over style**

- all of the style elements are handled through ui.R
- instead, you can create a www directory and then an index.html file in that directory
    - get rid of ui.R
    - go to [this link](http://rstudio.github.io/shiny/tutorial/#html-ui) for information on how to build the HTML file
- you just have to have specific js libraries and appropriately names ids and classes. This is outside the scope of this class
- for students with a lot of experience in html, js, css it would be a breeze and probably easier and more flexible than the R html controls in ui.R

**B. Other thing Shiny can do**

- allow users to upload or download files
- have tabbed main panels
- editable data tables
- have a dynamic UI
    - show new controls based on input values
- user-defined inputs and outputs
- put a submit button so Shiny only runs when user wants

**C. Distributing a Shiny app**

- when you run Shiny on your computer, you're running a local server and deploying locally. What if you want to distribute it?
- the quickest way is to send (or put on github or gist or dropbox or whatever) someone the app driectory and they can then call runApp
- you could create an R package and create a wrapper that calls runApp
    - of course, these solutions only work if the users knows R
- Another option is to run a shiny server
    - requires setting up a [Shiny server](http://www.rstudio.com/shiny/server/)
    - probably easiest if you use one of the virtual machines where they already have Shiny servers running well (for example, on AWS)
    - setting up a Shiny server is beyond the scope of this class as it involves some amount of linux server administration
    - groups are creating a Shiny hosting service that will presumably eventually be a fee for service or freemium service
    - BTW...don't put system calls in your code (this is one of the first things many of us do for fun, but it introduces security concerns)
        - don't call the Unix terminal or other stuff like that
        - don't run "generic server-side scripts" with R as the intermediary

####*Video 1.8 More advanced shiny discussion, reactivity*

**A. Details**

- code that you put before shinyServer in the server.R function gets called once when you do runApp()
- code inside the unnamed function of shinyServer(function(input,output)) but not in a reactive statement will run once for every new users (or page refresh)
- code in reactive functions of shinyServer get run repeatedly as needed when new values are entered (reactive functions are those like render*)

**B. Try it**

- hitting refresh will increment y but not entering a value in text box
- stuff outside the function call only gets updated once, regardless of page refreshes
- try ```runApp(display.mode='showcase')``` to see what Shiny is actually executing when it updates
- stuff inside the function only updated as needed (it doesn't just re-run all the commands every time you enter something)

####*Video 1.9 More advanced shiny, the reactive function*

**A. Reactive expressions**

- sometimes to speed up your app, you want reactive operations (those operations that depend on widget input values) to be performed outside of a render* statement
- for example, you want to do some code that gets reused in several render* statements and don't want to recalculate it for each
- the ```reactive()``` function is made for this purpose

**B. a server.R Example**

```{r react, eval=FALSE}
shinyServer(
    function(input, output){
        x <- reactive({as.numeric(input$text1)+100})
        output$text1 <- renderText({x()})  ## display input incremented by 100
        output$text2 <- renderText({x() + as.numeric(input$text2)}) ## display input incremented by 100 plus text 2
        }
    )
## now we only have to do the incrementing by 100 once, rather than incrementing by 100 a bunch of times
```

- really odd syntax for reactive variables
    - call x() to get value of the variable x
- a note: use CTRL + SHIFT + C to comment/uncomment quickly

####*Video 1.10 More advanced shiny, conditional e xecution of reactive statements*

**A. Non-reactive reactivity (what?)**

- sometimes you don't want shiny to immediately perform reactive calculations from widget inputs
    - you want the user to put all the stuff in before running
- in other words, you want something like a submit button
- some code to add a 'GO' button:

```{r gobutton, eval=FALSE}
## ui.R
shinyUI(pageWithSidebar(
    headerPanel("Hello Shiny!"),
    sidebarPanel(
        textInput(inputID="text1", label = "Input Text1"),
        textInput(inputId="text2", label = "Input Text2"),
        actionButton("goButton", "Go!")
        ),
    mainPanel(
        p('Output text1'),
        textOutput('text1'),
        p('Output text2'),
        textOutput('text2'),
        p('Output text3'),
        textOutput('text3')
        )
    ))

## server.R
shinyServer(
    function(input, output){
        output$text1 <- renderText({input$text1}) ## renders automatically (regardless of button)
        output$text2 <- renderText({input$text2})
        ## make rendering of text3 dependent on Go Button
        output$text3 <- renderText({
            input$goButton
            ## isolate this statement until button is pressed
            isolate(paste(input$text1, input$text2))
        })
    }   
)
```

**B. Discussion**

- notice it doesn't display output text3 until the go button is pressed
- input$goButton (or whatever you named it) gets increased by one for every time pushed
- so, when in reactive code (such as render or reactive) you can use conditional statements like below to only execute code on the first button press or to not execute code until the first or subsequent button press
    - ```if (input$goButton == 1){Conditional statements}```

####*Video 1.11 More advanced shiny, odds and ends*

**A. More on layouts**

- the sidebar layout with a main panel is the easiest
- using ShinyUI(fluidpage( is much more flexible and allows tighter access to the bootstrap styles
    - bootstrap style invented by twitter
- example [here](http://shiny.rstudio.com/articles/layout-guide.html)
- ```fluidRow``` statements create rows and then the ```column``` function from within it can create columns
- tabsets, navlists and navbars can be created for more complex apps

**B. Directly using html**

- for complex layouts, direct use of html is preferred
- if you know web development well, working with R for the layout is annoying
- some other notes on this page for people entering raw HTML (skip for now)

**C. Debugging techniques for Shiny**

- this can be difficult
- we saw that ```runApp(displayMode = 'showcase') highlights execution while a shiny app runs
- using ``cat`` in your code displays output to stdout (so R console)
- the broswer() function can interrupt execution and can be called conditionally [link](http://shiny.rstudio.com/articles/debugging.html)

####*Video 1.12 Manipulate*

**A. What is manipulate?**

- a tool in Rstudio for creating quick dynamic graphics
- suppose that you want to create a quick interactive graphic
    - you have to do it now
    - the intended users also use Rstudio
- ```manipulate`` is a really cool solution that might be all you need

**B. Documentation**

- Manipulate is well documented at the Rstudio website [here](http://www.rstudio.com/idea/docs/advanced/manipulate)
- from there, try this:
    - ```library(manipulate); manipulate(plot(1:x, x=slider(1,100)))```
    
**C. Example from the regression class**

```{r manip, eval=FALSE}
library(manipulate); library(usingR);
myHist <- function(mu){
    hist(galton$child, col="blue", breaks=100)
    lines(c(mu,mu), c(0,150), col="red", lwd=5)
    mse <- mean((galton$child-mu)^2)
    text(63, 150, paste("mu = ",mu))
    text(63, 140, paste("MSE = ",round(mse,2)))
}

## add slider
manipulate(myHist(mu), mu=slider(62, 74, step = 0.5))
```

####*Video 1.13 Intro to rCharts and GoogleVis*

**A. Interactivity in Graphs**

- give users some ability to manipulate
- rCharts, googleVis for interactive, engaging plots
    - make cool plots that don't require presentations

####*Video 1.14 rCharts introduction*

**A. rCharts**

- a way to create interactive js visualization using R
- So:
    - you don't have to learn complex tools, like D3
    - just work in R learning a minimal amount of new syntax
- rCharts was written by Ramnath Vaidyanathan (friend of the Data Science Series), who also wrote slidify, the framework we use for all of the lectures in the class
- this lecture is going through [the rCharts tutorial](http://ramnathv.github.io/rCharts/)
- flurry of StackExchange discussion about this

**B. An Example**

```{r rcharts, eval=TRUE, results= 'asis', comment = NA}
require(rCharts)
haireye = as.data.frame(HairEyeColor) 

## call nPlot for a specific js plotting library (assign the plot to a value)
n1 <- nPlot(Freq ~ Hair, group = 'Eye', type = 'multiBarChart',
            data = haireye[haireye$Sex=='Male',])

## show the plot
n1$show("iframesrc", cdn= TRUE)

## use this command (cat = print out an iframe) to embed into slidify
#cat(' <iframe src="fig/n1.html" width=100%, height=600></iframe>')

## to get this working, you need to call servr::httd() to open a test http session, then view the html output from knitr. giggity.
```

####*Video 1.15 rCharts more examples*

**A. Slidify interactive**

- the above was an example of embedding an rChart in a slidify document
    - in the YAML yaml ext_widgets : {rCharts: ["libraries/nvds3"]}
- or, if you use more than one library:
    - see notes
- google "slidify interactive" for tutorials

**B. Viewing the plot**

- the object n1 contains the plot
    - in RStudio, typing n1 brings up the plot in the RStudio viewer (or you can just not assign it to an object)
- do n1$ then hit TAB to see the various functions contained in the object
    - n1$html() prints out the html for the plot
- I do n1$save(filename) then bring the code back into slidify document
    - this is recommedned for slidify, but if you're just looking at the plot, it's unnecessary
    
**C. Another Example**

```{r examp2, results = 'asis'}
library(datasets); data(iris);
names(iris) = gsub("\\.","",names(iris))

r1 <- rPlot(SepalLength ~ SepalWidth | Species, data = iris, color = 'Species', type = 'point')

## show the plot
r1$show("iframesrc", cdn= TRUE)

##slidify stuff
# r1$save('fig/r1.html', cdn = TRUE)

# cat('<iframe src="fig/r1.html" width = 100%, height=600></iframe>')
```

- look on the internet and find lots of examples

**D. Example2 Facetted barplot**

```{r facets, results= 'asis'}
hair_eye <- as.data.frame(HairEyeColor)
r2 <- rPlot(Freq ~ Hair | Eye, color = 'Eye', data = hair_eye, type = 'bar')

## show the plot
r2$show("iframesrc", cdn= TRUE)

```

**E. How to get the hs/html or publish an rChart**

- now you can add whatever you'd like

```{r publish, eval=FALSE}
r1 <- rPlot(mpg ~ wt | am + vs, data = mtcars, type = "point", color = "gear")
r1$print("chart1") ## print out the js
r1$save('myPlot.html') ## save as html file
r1$publish('myPlot', host = 'gist') ## save to gist, rjson required
r1$publish('myPlot', host = 'rpubs') ## save to rpubs
```

**F. rCharts has links to several libraries**

- we'll do some more examples
- note Ramnath mentions that io2012 and polychart have conflicting js

**G. morris**

- time series plots with metadata!!!!

```{r morris, results='asis'}
library(rCharts)
data(economics, package="ggplot2")
econ <- transform(economics, data=as.character(date))
## make a line plot of two variable, private savings rate and unemployment(?)
m1 <- mPlot(x="date", y=c("psavert", "uempmed"), type = "Line", data = econ)
m1$set(pointSize = 0, lineWidth = 1)

## show the plot
m1$show("iframesrc", cdn= TRUE)

## slidify stuff
# m1$save('fi1/m1.html', cdn = TRUE)
# cat('<iframe src="fig/m1.html" width = 100%, height=600></iframe>')

```

**H. xCharts**

- kind of a stacked line graph

```{r xcharts1, results = 'asis'}
require(reshape2)
uspexp <- melt(USPersonalExpenditure)
names(uspexp)[1:2] <- c("category", "year")
x1 <- xPlot(value ~ year, group = "category", data = uspexp, type = "line-dotted")

## show the plot
x1$show("iframesrc", cdn= TRUE)

## slidify stuff
# x1$save('fig/x1.html', cdn=TRUE)
# cat('<iframe src = "fig/x1.html" width=100%, height=600></iframe>')
```
####*Video 1.16 rCharts mapping and discussion*

**A. Leaflet**

- this is a mapping library
- create an interactive map and add in pop-ups by long. and lat.

```{r leafletexample, results='asis'}
library(rCharts)
## open a leaflet plot
map3 <- Leaflet$new()

## set the view by this longitude and latitude. Include Zoom rate
## (don't worry, there is a zoom feature on the map)
map3$setView(c(51.505, -0.09), zoom = 13)

## add some markets (think Google Maps check, Apple pins)
## markers just need long., lat., and some text to show when you click them
map3$marker(c(51.5, -0.09), bindPopup = "<p> Hi. I am a popup.</p>")
map3$marker(c(51.495, -0.083), bindPopup = "<p> Hi. I am another popup </p>")

## show the plot
map3$show("iframesrc", cdn= TRUE)

## slidify stuff
# map3$save('fig/map3.html', cdn=TRUE)
# cat('Miframe src="fig/map3.html" width=100%, height=600></iframe>')
```

**B. Rickshaw**

- see the code later. Has a lisder for time, get stacked line charts

**C. highchart**

- does some cool stuff. Not sure what's going on. See code from lecture

**D. rCharts summarized**

- rCharts makes creating interactive javascript visualizations in R ridiculously easy
- However, non-trivial customization is going to require knowledge of javascript
- If what you want is not too big of a deviation from the rCharts examples, then it's awesome
    - otherwise, it's challenging to extend without fairly deep knowledge of the JS libraries that it's calling
- rCharts is under fairly rapid development
    - GoogleVis is pretty good too
- if you're using this with slidify, always have most up-to-date rCharts, slidify, RStudio
- delete old "assets" folders if you're having issues with the slidify-rCharts-RStudio combination

####*Video 1.17 GoogleVis*

**A. Google Vis API**

- make plots that apply in HTML pages
- see [this link](https://developers.google.com/chart/interactive/docs/gallery)
- interactive data visualizations
- the GoogleVis R function wraps these in R commands

**B. Basic Idea**

- The R function creates an HTML page
- the HTML page calls Google Charts
- the result is an interactive HTML graphic
- [googleVis documentation](http://cran.r-project.org/web/packages/googleVis/vignettes/googleVis.pdf)

**C. Motion Charts!!!**

```{r googvis1, eval=FALSE}
suppressPackageStartupMessages(library(googleVis))
M <- gvisMotionChart(Fruits, "Fruit", "Year", options = list(width=600, height = 400))

## show in browser:
plot(M)

## slidify
# print(M, "chart")
```

**D. Charts in googleVis**

- Motion charts: gvisMotionChart
- Interactive maps: gvisGeoChart
- Interactive tables: gvisTable
- Line charts: gvisLineChart
- Bar charts: gvisColumnChart
- Tree maps: gvisTreeMap


**E. Plots on maps**

```{r geochart, eval=FALSE}
G <- gvisGeoChart(Exports, locationvar = "Country", colorvar = "Profit", options = list(width = 600, height = 400))

## see in browser:
plot(G)

## print for slidify
# print(G, "chart")
```

- another geoplot, but with only a specific region

```{r europe, eval=FALSE}
## same plot, but only European countries
G2 <- gvisGeoChart(Exports, locationvar = "Country", colorvar = "Profit", options = list(width = 600, height = 400, region="150"))

## see in browser:
plot(G2)

## print for slidify
# print(G2, "chart")
```

- finding options for googlevis:
    - go to the "Configuration options" on documentation
- [the gVis documentation](https://developers.google.com/chart/interactive/docs/gallery/geochart)

**F. Setting More Options**

- from [a talk](https://github.com/mages/Introduction_to_google_Vis/blob/gh-pages/index.Rmd) by googleVis creater
- see the notes here (way too much code to type)
- lots of customizability
    - you need to do some digging into the googleVis API
    
**G. Combining Mulltiple Plots Together (multi-panel)**

```{r panels,eval=FALSE}
library(googleVis)
G <- gvisGeoChart(Exports, "Country", "Profit", options = list(width = 200, height = 100))
T1 <- gvisTable(Exports, options = list(width = 200, height = 270))
M <- gvisMotionChart(Fruits, "Fruit", "Year", options = list(width = 400, height = 370))

## do merges two at a time

## first merge the map and table
GT <- gvisMerge(G, T1, horizontal = FALSE)

## then merge this merged thing with the motion chart
GTM <- gvisMerge(GT, M, horizontal = TRUE, tableOptions = "bgcolor=\"#CCCCCC\" cellspacing = 10")

## check it out:
plot(GTM)
```

- note in the code above how the widths and heights were set to match up nicely
- this is just gorgeous
- if you use print(object_name), it will give you the raw HTML code
    - you could look at how this works and directly edit the code if you want
    
**Things you can do with GoogleVis**

- the visualization can be embedded in websites with HTML code
- dynamic visualization can be built with Shiny, Rook, and R.rsp
- embed them in R markdown based documents
    - set results = 'asis' in the chunk options
    - can be used with knitr and slidify
- you can type demo(googleVis) to get a plot by plot walkthrough
- check out the googleVis faq's and gallery (online)

####*Video 1.18 shinyApps.io*

**A. How to get your ShinyApps.io account**

- to deploy your app, you can use deployAPP('path/to/your/app') to put your stuff on ShinyApps.io
- you can view your apps from in there

####*Video 1.19 plotly*

**A. What is plotly?**

- plotly is a platform for creating interactive charts and doing analysis
    - focus here is on plotly as it applies to R
- interaction with specific web platform

**B. plotly example**

```{r plotly, eval=FALSE}
## load data
load("courseData.rda")

## libraries
library(plotly); library(ggplot2)

## make a plot
g <- ggplot(myData, aes(y=enrollment, x=class, fill=offering))
g <- g + geom_bar(stat = "identity")
g

## get it into plot.ly
py <- plotly()
out <- py$ggplotly(g) ## this will open up a browser window the the plot in it
out$response$url ## this is the url of the plotly plot
    ## the plotly plot will be interactive!!
```

- once the plot is up there, use "save and edit" to change a bunch of features
- this is a "social graphing platform"
    - share your graph on facebook and twitter