---
title: "Analytic Edge Kaggle NYT classifiction"
author: "Ryan Zhang"
date: "Thursday, April 16, 2015"
output:
  html_document:
    fig_height: 4
    fig_width: 5
    highlight: espresso
    keep_md: yes
    theme: cosmo
    toc: yes
---

#0 Environment   
##0-1 Set working directory     
```{r }
setwd("~/GitHub/edX/The Analytic Edge/Kaggle")
```

##0-2 Load Libraries   
```{r warning=FALSE,message=FALSE}
library(tm)
library(e1071)
library(neuralnet)
library(randomForest)
library(ROCR)
library(party)
library(rCUR)
```

##0-3 Function Definition     
 table     
```{r}
tCor <- function(t)round(t[,2]/rowSums(t),2)*100 
```
 dummy encoding 
```{r}
dummyEncoding <- function(df, colname){
  dDF <- as.data.frame(model.matrix(~df[,colname]))
  names(dDF) <- paste(colname,as.character(levels(df[,colname])),sep="")
  dDF}
```

#1 Data Preparing  ׼ 
##1-1 Loading װ 
```{r}
NewsTrain <- read.csv("NYTimesBlogTrain.csv", stringsAsFactors = F)
NewsTest <- read.csv("NYTimesBlogTest.csv", stringsAsFactors = F)
```

##1-2 Ԥ 
Store the number of training data points and the number of testing data points.        
 ¼һ ѵ  ݺͲ ݵ  
```{r}
ntrain <- nrow(NewsTrain)
ntest <- nrow(NewsTest)
ntrain
ntest
```

"Popular"" is the dependant variable, store it in a separate vector "Y", and delete the colomn from the dataframe "NewsTrain".      
 Popular 
```{r}
Y <- as.factor(NewsTrain$Popular)
NewsTrain$Popular <- NULL
```

Combine "NewsTrain" and "NewsTest" into a single dataframe for the purpose of data preparing      
    
**
```{r}
OriginalDF <- rbind(NewsTrain, NewsTest)
```

Filling empty entries for the first three columns with name "Other"      
 ǰ ġ á Other  
```{r}
for (i in 1:nrow(OriginalDF)){
  for (j in 1:3){
    if (OriginalDF[i,j] == ""){
      OriginalDF[i,j] <- "Other"}}}
```

Change the first three columns to be factors     
 ǰ  ĳ factor 
```{r}
OriginalDF$NewsDesk <- as.factor(OriginalDF$NewsDesk)
OriginalDF$SectionName <- as.factor(OriginalDF$SectionName)
OriginalDF$SubsectionName <- as.factor(OriginalDF$SubsectionName)
```

Log Transform "WordCount"      
 WordCount ת ת    ̬ ֲ    
```{r}
# OriginalDF$ZWordCount <- with(OriginalDF, (WordCount - mean(WordCount))/sd(WordCount))
OriginalDF$NWordCount <- log(OriginalDF$WordCount + 1)
```

Conver the PubDate and time variable to be more R friendly and extract the hour of day, the day on month and the day of week to be seperate variables. Finally delete the PubDate column.       
 PubDate ĳ R -ʱ  ʽ  ܼ  ÿ ¼ Լ ÿ 켸 Щ Ϣ ȡ ɾ ԭ PubDate
```{r}
OriginalDF$PubDate <- strptime(OriginalDF$PubDate, "%Y-%m-%d %H:%M:%S")
OriginalDF$Hour <- as.factor(OriginalDF$PubDate$h)
OriginalDF$Wday <- as.factor(OriginalDF$PubDate$wday)
OriginalDF$Mday <- as.factor(OriginalDF$PubDate$mday)
# OriginalDF$isWeekend <- as.numeric(OriginalDF$Wday %in% c(0,6))
OriginalDF$PubDate <- NULL
```

Generate training and testing set    
 ѵ  Ͳ    
```{r}
train <- OriginalDF[1:ntrain, c(1:3,9:12)]
test <- OriginalDF[(ntrain+1):nrow(OriginalDF),c(1:3,9:12)]
```

##2 Exploratory Data Analysis  ̽ ʽ  ݷ      
First Explore the few factor variable and their relationship to the depandent variable.    
 ȿ  ǰ factor ҪԤ  Popular֮ Ĺ ϵ     
```{r}
tNewsDesk <- table(OriginalDF$NewsDesk[1:ntrain], Y)
t(tNewsDesk)
tCor(tNewsDesk)
plot(tCor(tNewsDesk))

tSectionName <- table(OriginalDF$SectionName[1:ntrain], Y)
t(tSectionName)
tCor(tSectionName)
plot(tCor(tSectionName))

tSubsectionName <- table(OriginalDF$SubsectionName[1:ntrain], Y)
t(tSubsectionName)
tCor(tSubsectionName)
plot(tCor(tSubsectionName))
```
    
SectionName, SubsectionName,NewsDeskӦ ö  Ԥ  ı  Ӧ ñ     

Looking at the text contents    
  ı  Ϣ   
It seems that the "Snippet" is almost redudent with "Abstract", in since 98% cases they are the same. And "Abstract" contains a little bit more infomation than "Snippet"      
SnippetӦ ú Abstract  غ ݷǳ ࣬ǰ òT ƶ ں ߣ ֻ ú ߾ͺ ˡ     
```{r}
sum(OriginalDF$Snippet == OriginalDF$Abstract)/nrow(OriginalDF)
which(OriginalDF$Snippet != OriginalDF$Abstract)[1]
OriginalDF[22,5]
OriginalDF[22,6]
```

Looking at WordCount    
    
The distribution of WordCount seems to be a longtail / power-law distribution.    
  ķֲ ƺ ɷֲ      
```{r}
summary(OriginalDF$WordCount)
hist(OriginalDF$WordCount, breaks = 70)
hist(OriginalDF$NWordCount)
```

Looking at publication day/weekday/hour related to Popular   
 Сʱ  ܼ  ÿ ¼ ţ  Щ û    
```{r}
tHour <- table(OriginalDF$Hour[1:ntrain] , Y)
t(tCor(tHour))
plot(tCor(tHour))

tWday <- table(OriginalDF$Wday[1:ntrain], Y)
t(tCor(tWday))
plot(tCor(tWday))

tMday <- table(OriginalDF$Mday[1:ntrain], Y)
t(tCor(tMday))
plot(tCor(tMday))

#tWeekend <- table(OriginalDF$isWeekend[1:ntrain], Y)
#tCor(tWeekend)
#plot(tCor(tWeekend))
```
ÿ ¼ ſ  ȥûɶ    

#3 Model fitting  ģ    
##3-1 randomForest on Non-Text Features  ò ı  ȡ   ɭ 
randomForest model    
  ɭ ģ     
public board auc = 0.92699
```{r  cache=TRUE}
set.seed(880306)
rfModel <- randomForest(x = train,
                        y = Y,
                        ntree = 1000,
                        mtry = 2,
                        nodesize = 4,
                        importance = F,
                        proximity = F)
```

Evaluate on Training set 
  ݲ Լ  ģ 
```{r cache=TRUE}
rfPred <- predict(rfModel, train, type = "prob")
table(rfPred[,2] > 0.5,Y)

prediction <- ROCR::prediction(rfPred[,2], Y)
perf <- performance(prediction, "tpr", "fpr")
plot(perf, colorize = T, lwd = 2)
auc <- performance(prediction, "auc")
auc@y.values
```

Make prediction with randomForest model
  ɭ ģ Ԥ 
```{r }
tpred <- predict(rfModel, test, type = "prob")
MySubmission = data.frame(UniqueID = NewsTest$UniqueID, Probability1 = tpred[,1])
write.csv(MySubmission, "rfRegularFeatures.csv", row.names = F)
```

#4 Try feature engineering with text content  ͨ  ı      
##4-1 TFIDF   
Extract all headline and abstract to form a corpus    
   
```{r cache=TRUE}
text <- vector()
for (i in 1:nrow(OriginalDF)) {
  text <- rbind(text, paste(OriginalDF$Headline[i], " ", OriginalDF$Abstract[i]))
}

Corpus <- Corpus(VectorSource(text))
```

Standard Corpus processing     
 ׼  Ͽ⴦      
```{r cache=TRUE}
Corpus <- tm_map(Corpus, tolower)     
Corpus <- tm_map(Corpus, PlainTextDocument)    
Corpus <- tm_map(Corpus, removePunctuation)    
Corpus <- tm_map(Corpus, removeWords, stopwords("english"))     
Corpus <- tm_map(Corpus, stemDocument)    
```

Document ~ TF-IDF matrix    
  ĵ ~TFIDF      
```{r cache=TRUE}
dtm <- DocumentTermMatrix(Corpus, control = list(weighting = weightTfIdf))    
```

Get the terms    
 ȡ  б      
```{r }
terms <- dtm$dimnames$Terms    
terms[5101:5110]
```

Get the matrix for training and testing set     
 ֱ  ѵ  Ͳ ݵ Document~TF-IDF      
```{r }
dtmTrain <- dtm[1:ntrain,]
dtmTest <- dtm[(1+ntrain):dtm$nrow,]
```

Get frequent terms matrix for testing set
 ò Լ  Ƶ 
```{r }
sparseTest <- removeSparseTerms(dtmTest, 0.96)
wordsTest <- as.data.frame(as.matrix(sparseTest))
termsTest <- names(wordsTest)
```

Filter the dtm based on frequent terms in testing set    
  ݲ Լ  Ƶ   ԭ  ľ ɸѡ     
```{r cache=TRUE}
cols <- vector()
for (i in 1:length(termsTest)){
  cols = c(cols, which((terms == termsTest[i]) == T))}
dtmFiltered <- dtm[,cols]
```

Text Feature    
 ı        
```{r cache=TRUE}
termFeatures <- as.data.frame(as.matrix(dtmFiltered))
row.names(termFeatures) <- c(1:nrow(OriginalDF))
```

Append text features to the dataframe    
```{r}
TextADDDF <- as.data.frame(termFeatures)
```

```{r}
tatrain <- cbind(train, TextADDDF[1:ntrain,])
tatest <- cbind(test, TextADDDF[(ntrain+1):nrow(TextADDDF),])
```

##4-2 randomForest model with text features added    ı  ɭ ģ   
public board auc = 
```{r cache=TRUE}
set.seed(880306)
tarfModel <- randomForest(x = tatrain,
                        y = Y,
                        ntree = 1000,
                        nodesize = 4,
                        importance = T,
                        proximity = F)
```

Look at the importance of features via training randomForest   
  ı  Ҫ     
```{r}
t(tarfModel$importance[,4])
```
  
ò ƶ ͦ ͵ġ      

Make prediction on the training set
 ü ı  ɭ ģ Ͷ ѵ  ݽ  Ԥ 
```{r cache=TRUE}
tarfPred <- predict(tarfModel, tatrain, type = "prob")
table(tarfPred[,2] > 0.5,Y)

prediction <- ROCR::prediction(tarfPred[,2], Y)
perf <- performance(prediction, "tpr", "fpr")
plot(perf, colorize = T, lwd = 2)
auc <- performance(prediction, "auc")
auc@y.values
```

Make prediction with randomForest model
  ı  ɭ ģ Ԥ 
```{r}
tatpred <- predict(tarfModel, tatest, type = "prob")
MySubmission = data.frame(UniqueID = NewsTest$UniqueID, Probability1 = tatpred[,1])
write.csv(MySubmission, "rfText.csv", row.names = F)
```

##4-3 Why not a neural net?        
Neural net based on numerical features.    
 ֵ   磨û GPU        
```{r cache=TRUE}
nntrain <- tatrain[,c(4,8:26)]
nntest <- tatest[,c(4,8:26)]
nntrain$Popular <- as.numeric(as.character(Y))
nameofvars <- names(nntrain)
nnformula <- as.formula(paste("Popular ~", 
                        paste(nameofvars[!nameofvars %in% "Popular"], collapse = " + ")))
ptm <- proc.time()
nn <- neuralnet(formula = nnformula,
                data = nntrain, 
                hidden = 2,
                threshold = 0.02,
                stepmax = 1e8,
                err.fct="ce",
                linear.output=FALSE)
plot(nn)
pnn <- compute(nn,nntrain[,1:20])
summary(pnn$net.result)
nnpredict <- as.vector(pnn$net.result)
prediction <- ROCR::prediction(nnpredict, Y)
perf <- performance(prediction, "tpr", "fpr")
plot(perf, colorize = T, lwd = 2)
auc <- performance(prediction, "auc")
auc@y.values
proc.time() - ptm
```

##4-4 Support Vector Machine ֧ 
 Dummy Encoding
```{r cache=TRUE}
taDF <- rbind(tatrain, tatest)
names(taDF)
tadDF <- taDF[,c(4,8:26)]
toEnco <- c("NewsDesk","SectionName","SubsectionName","Hour","Wday","Mday")
for (colname in toEnco){
  tadDF <- cbind(tadDF, dummyEncoding(taDF, colname))
}
names(tadDF) <- make.names(tadDF)
tadtrain <- tadDF[1:ntrain,]
tadtest <- tadDF[((1+ntrain):nrow(tadDF)),]
```

SVM Model Fitting ֧ 
```{r cache=TRUE, Warning=FALSE}
svmModel <- svm(x = tadtrain, 
                y = Y,
                cost = 100, gamma = 1, probability = T)
svmpred <- predict(svmModel, tadtrain, decision.values = T, probability = T)
svmpredp <- attr(svmpred, "probabilities")[,1]
prediction <- ROCR::prediction(svmpredp, Y)
perf <- performance(prediction, "tpr", "fpr")
plot(perf, colorize = T, lwd = 2)
auc <- performance(prediction, "auc")
auc@y.values
svmtpred <- predict(svmModel, tadtest, probability = T)
svmtpredp <- attr(svmtpred, "probabilities")[,1]
MySubmission = data.frame(UniqueID = NewsTest$UniqueID, Probability1 = svmtpredp)
write.csv(MySubmission, "svmText.csv", row.names = F)
```

##4-5 Simple Ensemble  򵥵 ģ 
public board auc 0.91463
```{r cache=TRUE}
enpred <- 0.5*svmtpredp+0.5*tatpred[,2]
MySubmission = data.frame(UniqueID = NewsTest$UniqueID, Probability1 = enpred)
write.csv(MySubmission, "svm_rf_ensemble.csv", row.names = F)
```


#5 Clustering on TFIDF matrix   ࿴ 
```{r cache=TRUE}
cTest <- removeSparseTerms(dtmTest, 0.98)
cWords <- as.data.frame(as.matrix(cTest))
cTerms <- names(cWords)

cols <- vector()
for (i in 1:length(cTerms)){
  cols = c(cols, which((terms == cTerms[i]) == T))
}
cdtm <- dtm[,cols]
cMatrix <- as.matrix(cdtm)
cDF <- as.data.frame(cMatrix)

kmclusters <- kmeans(cDF, 9, iter.max = 5000)

tactrain <- cbind(tatrain, kmclusters$cluster[1:ntrain])
names(tactrain) <- c(names(tactrain)[1:26],"cluster")
tactest <- cbind(tatest, kmclusters$cluster[(1+ntrain):nrow(OriginalDF)])
names(tactest) <- c(names(tactest)[1:26],"cluster")
```

##5-1 Another randomForest with cluster labels added   Ͼ ǩ һ  ɭ  
public board auc = 0.92495
```{r cache=TRUE}
set.seed(1126)
tacrfModel <- randomForest(x = tactrain,
                        y = Y,
                        ntree = 1000,
                        mtry = 6,
                        nodesize = 4,
                        importance = T,
                        proximity = F)
t(tacrfModel$importance[,4])
tacrfPred <- predict(tacrfModel, tactrain, type = "prob")
table(tacrfPred[,2] > 0.5,Y)

prediction <- ROCR::prediction(tacrfPred[,2], Y)
perf <- performance(prediction, "tpr", "fpr")
plot(perf, colorize = T, lwd = 2)
auc <- performance(prediction, "auc")
auc@y.values

tacrfPred <- predict(tacrfModel, newdata = tactest, type = "prob")
MySubmission = data.frame(UniqueID = NewsTest$UniqueID, Probability1 = tacrfPred[,2])
write.csv(MySubmission, "rfTextCluster.csv", row.names = F)
```

#6 Matrix Factorization  ֽ   
##6-1 SVD  ֵ ֽ 
```{r cache=TRUE}
s <- svd(cMatrix)
Sig <- diag(s$d)
plot(s$d)
totaleng <- sum(Sig^2)
engsum <- 0
for (i in 1:nrow(Sig)){
  engsum <- engsum + Sig[i,i]^2
  if (engsum/totaleng > 0.8){
    print(i)
    break}}
```
 Ҫ60 ܱ  ԭTF-IDF 80%  

##6-2 CUR Matrix Decomposition CUR ֽ 
```{r cache=TRUE}
res <- CUR(cMatrix, c = 10, r = 84, k = 60)
```

 ͶӰ ӵ ѵ  
```{r cache=TRUE}
ncolC <- ncol(getC(res))
Ak <- getC(res) %*% getU(res)[,1:ncolC]
AkDF <- as.data.frame(Ak)
tacftrain <- cbind(tactrain, AkDF[1:ntrain,])
tacftest <- cbind(tactest, AkDF[((1+ntrain):nrow(AkDF)),])
```

##6-3 randomForest with pc added  Ҫ ɷ ͶӰ ɭ 
public board auc = 0.92511
```{r cache=TRUE}
set.seed(1126)
tacfrfModel <- randomForest(x = tacftrain,
                        y = Y,
                        ntree = 1000,
                        mtry = 8,
                        nodesize = 4,
                        importance = T,
                        proximity = F)
t(tacfrfModel$importance[,4])
tacfrfPred <- predict(tacfrfModel, tacftrain, type = "prob")
table(tacfrfPred[,2] > 0.5,Y)

prediction <- ROCR::prediction(tacfrfPred[,2], Y)
perf <- performance(prediction, "tpr", "fpr")
plot(perf, colorize = T, lwd = 2)
auc <- performance(prediction, "auc")
auc@y.values

tacfrfPred <- predict(tacfrfModel, newdata = tacftest, type = "prob")
MySubmission = data.frame(UniqueID = NewsTest$UniqueID, Probability1 = tacfrfPred[,2])
write.csv(MySubmission, "rfTextClusterFactorization.csv", row.names = F)
```

