<center> <h2>Principles of Computing - Week 4</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 4.1 - The Importance of Combinatorics*

**A. Notes**

- taking a bunch of discrete shapes and putting them together
- "shortest path" types of questions
- combinatorics:
    - we can search a sample space to figure out how hard the problem is
- brute-force:
    - try all the combinations
- you can't say math is more important than programming or vice versa

####*Video 4.2 - Enumeration*

**A. Notes**

- combinatorics = studying taking a set of items and combining them to make more complicated objects

```{r eval=FALSE}
"""
Functions to enumerate sequences of outcomes
Repetition of outcomes is allowed
"""


def gen_all_sequences(outcomes, length):
    """
    Iterative function that enumerates the set of all sequences of
    outcomes of given length
    """
    
    ans = set([()])
    for dummy_idx in range(length):
        temp = set()
        for seq in ans:
            for item in outcomes:
                new_seq = list(seq)
                new_seq.append(item)
                temp.add(tuple(new_seq))
        ans = temp
    return ans


# example for digits
def run_example1():
    """
    Example of all sequences
    """
    outcomes = set([0, 1, 2, 3, 4, 5, 6, 7, 8, 9])
    #outcomes = set(["Red", "Green", "Blue"])
    #outcomes = set(["Sunday", "Mondy", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"])
    
    length = 2
    seq_outcomes = gen_all_sequences(outcomes, length)
    print "Computed", len(seq_outcomes), "sequences of", str(length), "outcomes"
    print "Sequences were", seq_outcomes


run_example1()



def gen_sorted_sequences(outcomes, length):
    """
    Function that creates all sorted sequences via gen_all_sequences
    """    
    all_sequences = gen_all_sequences(outcomes, length)
    sorted_sequences = [tuple(sorted(sequence)) for sequence in all_sequences]
    return set(sorted_sequences)


def run_example2():
    """
    Examples of sorted sequences of outcomes
    """
    # example for digits
    outcomes = set([0, 1, 2, 3, 4, 5, 6, 7, 8, 9])
    #outcomes = set(["Red", "Green", "Blue"])
    #outcomes = set(["Sunday", "Mondy", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"])
    
    length = 2
    seq_outcomes = gen_sorted_sequences(outcomes, length)
    print "Computed", len(seq_outcomes), "sorted sequences of", str(length) ,"outcomes"
    print "Sequences were", seq_outcomes
    
run_example2()
```

- think of reducing to combinations as "compression" of the possible space

####*Video 4.3 - Permutations and Combinations*

**A. Notes**

- when we enumerate, we allow repeated outcomes
- when we compute permutations, we don't
- remember that the empty sequence is a permutatation of length 0
- permutations:
    - order matters and no repeats
- combinations:
    - order is irrelevant but no repeats allowed
- Python is able to handle arbitrarily large integers
- there will always be less combinations than permutation

####*Video 4.4 - Combinatorics and Password Breaking*

**A. Notes**

- counting allows us to estimate running time, size of sample space
- 2^10 = 1024
- increase the outcome space by enriching the set or doing more trials
- why put letter and number constraints on passwords?
    - not about counting, about predictability
    - humans use mostly alphas
    - constraints force you to use passwords that are harder to guess
    - "humans are not random password generators'

####*Video 4.5 - Debugging*

**A. Notes**

- debuggin can be fun!
- patience is key
- be thoughtful about what you're doing and what the code should be doing
- modularity is key:
    - most of the battle is just finding the bug
- a lot of debugging is about confirming that the code is doing what you think it is
- "it seems to run" doesn't mean your program is bug-free!
- confirm the truth, try to narrow down the bug's location

####*Video 4.6 - Yahtzee*

**A. Notes**

- in our version, you win with the sum of all numbers of the same type
    - like just playing the top part
- your goal here is to create a computer player
- make a strategy function to tell you what to do on the last roll of Yahtzee

####*Homework*

Problem: In this week's lectures, we discussed an iterative approach to generating all sequences of outcomes where repeated outcomes were allowed. Starting from this program template, implement a function gen_permutations(outcomes, num_trials) that takes a list of outcomes and a number of trials and returns a list of all possible permutations of length num_trials from this set of outcomes.
Hint: gen_permutations can be built from gen_all_sequences by adding a single if statement that prevents repeated outcomes. When you believe that your code works correctly, select the answer printed at the bottom of the console.

Solution:

```{r eval=FALSE}
"""
Function to generate permutations of outcomes
Repetition of outcomes not allowed
"""

def gen_permutations(outcomes, length):
    """
    Iterative function that generates set of permutations of
    outcomes of length num_trials
    No repeated outcomes allowed
    """
    
    ans = set([()])
    
    for dummy_idx in range(length):
        temp = set()
        print ans
        for seq in ans:
            for item in outcomes:
                new_seq = list(seq)
                #print "New seq", new_seq
                #print item
                if len(new_seq) > 0: 
                    if not (item in new_seq):
                        new_seq.append(item)
                        #print "After: ", new_seq
                        temp.add(tuple(new_seq))
                else:
                    new_seq.append(item)
                    temp.add(tuple(new_seq))
            
        ans = temp
    return ans

def run_example():

    # example for digits
    outcome = set(["a", "b", "c", "d", "e", "f"])
    
    permutations = gen_permutations(outcome, 4)
    print "Computed", len(permutations), "permutations of length", str(4)
    print "Permutations were", permutations
    
    permutation_list = list(permutations)
    permutation_list.sort()
    print
    print "Answer is", permutation_list[100]

    

run_example()


```

