<center> <h2>Developing Data Products - Week 2</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 2.1 Presenting Data Analysis Writing a Data Report*

**A. A Data Analysis Report**

- the most common type of data product
- prompt file:
    - explain the problem at hand
- Data folders:
    - located all the data in a single folder
    - keep track of where data came from
- raw folders
    - raw code folder, final code folder
    - final code...only the stuff you want to report/share
- final report:
    - text, conclusion, references, figures, figure caption

####*Video 2.2 Slidify Intro*

**A. What is slidify?**

- a way to create data-centric presentations
- slidify is an almagamtion of other technologies including knitr, Markdown, and several javascript libraries for HTML5 presentations
- Slidify is infinitely extendable and customizable, yet easy to use!
- Slidify allows embedded code chunks and mathematical formulas which keeps your presentation reproducible
    - type out your math in LaTex
    - renders with jx MathJax
- slidify presentations are just HTML files, so you can view them with any web browser and share them easily on Github, Dropbox, or your own website

**B. Getting started withSlidify**

- set the working directory to where you want to create your slidify project

```{r slide1, eval=FALSE}
setwd('C:\\Users\\James\\Documents\\MOOCS\\MOOC_Stuff\\Developing_Data_Products')
```

- create your project and give your project a name (my project is named "first_deck")

```{r slidify2, eval=FALSE}
author("first_deck")
```

- this causes the following:
    1. A directory with the name of yoru project is created inside of yoru current directory
    2. Inside of this directory an assets directory and a file called "index.RMD" is created
    3. The assets directory is populated with the following empty folders: css, img, js, and layouts
    4. The newly created index.Rmd R Markdown file open up in RStudio
    - Any custom css, images, or javascript you want to use should respectively be put into the newly created css, img, and js folders
    
####*Video 2.3 Slidify working it out*

**A. Getting to know the YAML part**

```{r getitgoing, eval=FALSE}
author("first_deck")
slidify('index.Rmd') ## knit it
browseUrl('index.html') ## view the deck
```



####*Video 2.4 Slidify customization*

**A. Getting to know index.Rmd: YAML**

- index.Rmd is the R markdown document which you will use to compose
- the first part of an index.Rmd file is a bit of YAML code which will look like this
- framework - formatting of stuff
- highlighter - effects
- hitheme - different themes
- widgets - a couple of styles, widgets (math, quiz, others)
- mode - tell it where and when you want to give presentation (depends whether or not you have internet)
    - libraries might be copied to local if you want a standalone file
- You can edit your YAML to include the title, subtitle, author, and job of the author, including what slide framework you wish to use, which code highlighter you want to use, and any widgets you want to include
- other fields you can include in your YAML:
    - a log to appear in your title slude under ```logo```
    - the path to your assets folders and the paths to any other folders you may be using under ```url```
    - the specific theme for your code highlighter of choice under ```hitheme```
    
**B. Making Slides**

- whater you put after "##" will be the title of the slide
- ---marks the end of the slide
- .class #id are CSS attributes you can use to customize the slide
- whatever you put between ### and --- is upt to you...as long as it is valid R Markdown or HTML

####*Video 2.5 Slidify more details*

**A. Makings Slides (cont.)**

- to compile yoru presentation make sure the wd contains you index.Rmd file and enter ```slidify('index.Rmd')``

**B. Publishing to Github**

- first, log in to GitHub and create a new empty repo
- use the following command, but replace user with your username and repo with the name of your new repository (both arguments are strings)
- ```publish_github(user, repo)```

**C. HTML 5 Deck Frameworks**

- see notes

**D. Mathjax**

- You can include LaTeX math formatting as follows
- edit your YAML widgets: [mathJax]

**E. More Advanced HTML**

- just put the html right in the markdown file
- especially useful for stuff like images or tables
    - you can always edit the final html slide
    - this isn't the best solution...why not just code html right from the beginning?
- similarly, you can incorporate JS or anything else you want

**F. Adding interactive elements to slidify**

- you can add interactive elements to slidify:
    - quiz questions
    - interactive Rcharts plots
    - Shiny apps
- of course, you could do this directly with html/js
- more easily, the dev version of slidify has this built in
    - see http://slidify.github,io/dcmeetup/demos/interactive
**G. A Sample Quiz Question**

## Question 1

What is 1 +1?

1. 1
2. _2_
3. 3
4. 4

*** .hint this is a hint
*** .explanation This is an explanation

---

####*Video 2.6 Slidify reminder about knitting R*

####*Video 2.7 RStudio Presenter 1 Introduction and getting started*

**A. What is this?**

- RStudio created a presentation authoring tool within their development environment
- If you are familiar with slidify, you will also be familiar with this tool
    - code is authored in a generalized markdown format that allows for code chunks
    - the output is an html5 presentation
    - the file index for the presenter file is .Rpres, which gets converted to an .md file and to an html file if desired
    - there's a preview tool in RStudio and GUIs for publishing to Rpubs or viewing/creating an html file
    
**B. Authoring Content**

- [a great guide](http://www.rstudio.com/ide/docs/presentations/overview)
- Quick start:
    - ```file``` then ```New File``` then R Presentation
    - use basically the same R markdown format for authoring as slidify/knitry
- view your stuff in "presentations"
you can click the question mark --> "Markdown Quick Reference" to get some MD formatting help

####*Video 2.8 RStudio Presenter 2 Authoring details*

**A. Compiling and tools**

- RStudio auto formats and runs the code when you save the document
- Mathjax JS library is loaded by default so you can use LaTeX
- slide navigation button on the preview; clicking on the notepad icon takes you to that slide in the deck
- clicking on more yields options
    - clearing the knitr cache
    - viewing in a broswer (creates a temporary html file in AppData/local/temp)
    - create a html file to save where you want
- a refresh button

**B. Visuals**

- RStudio has made it easy to get some cool html5 effects, like cube transitions with simple options in YAML-like code after the first slide
    - such as ```transition: rotate```
- link to transition options on rstudio.com

**C. Hierarchical Organization**

- if you want a hierarchical organization structure, just add a ```type: typename`` option after the slide
    - create slide style for that level
- see that wwww.rstudio.com/ide/docs/presentations ... tutorial and notes
- split up two columns by putting "***" on a line by itself with blank lines before and after

**D. Changing the slide font**

- ```font-family: fontname``` option
- you could even use ```font-import: url``` to import fonts
- important caveats
    - fonts must be present on the system that you're presenting or
    - you have to be connected to the internet to use an imported font (don't do this for offline presentation
- check out: http://fonts.googleapis.com/css?family=Risque

**E. Really changing things**

- If you know html5 and CSS well, you can basically change whatever you want
- a css file with the same names as your presentation will be autoimported
- you can use ```css: file.css``` to import a css file
- you have to create named classes and then use ```class: classname``` to get slide-specific style control from your css
    - or you can apply within a <span>
- Ultimately, you have an html file, which you can edit as you wish
    - this should be viewed as a last resort, as the whole point is to have reproducible presentations, but may be the easiest way to get the exact style control you want for a final product


####*Video 2.9 RStudio Presenter 3 DIscussion and comparison with Slidify*

**A. Slidify vs. R Studio Presenter**

- slidify has more options and styles
- slidify more fleixible from RMD file
    - large user base too
    - much steeper learning curve
- RStudio Presenter is more GUI oriented
    - really easy to get started
    - smaller set of easy styles and options
    - default styles look very nice
    - ultimately as flexible as slidify with a little CSS and HTML knowledge

####*Video 2.10 Very quick introduction to gh-pages*

**A. Hosting Stuff on gh-pages**

- initialize a git repo
- add you presentations and stuff to it
- commit the files to the repo
- push them to a github repo
- create a branch:
    - ```git branch gh-pages```
    - ```git checkout gh-pages```
    - ```git push origin gh-pages```
- switch branches on github to gh-pages (click and pick)
    - ```touch .nojekyll```
    - this file tells git when it's hosting to not do fancy html stuff
    - ```git add .nojekyll```
    - commit and push that change
    - ```git push origin gh-pages```
- go to username.github.io/reponame/htmlfile