---
output: pdf_document
fontsize: 10.5pt
geometry: margin=0.5in
citecolor: cyan
---

<center> <h1>Statistical Inference Final Project Part 2: ToothGrowth</h1> </center>
<center> By [James Lamb](http://www.linkedin.com/in/jameslamb1/)</center>
\hrulefill

\section{I. Introduction}

In this brief report, I present a basic summary of the ToothGrowth dataset available in the ```datasets``` package in R. I use basic plotting, R summary functions, confidence intervals and hypothesis tests to describe some characteristics of the 60-observation dataset.

\section{II. Exploratory Analysis}

The code to load the data is:

```{r}
library(datasets)
data(ToothGrowth)
```

Figure 1 in the Appendix shows some preliminary evidence that guinea pigs who received higher doses of Vitamin C had higher rates of tooth growth. There is no overlap in the distributions of tooth length by dose, providing strong evidence that the treatment does have a impact on tooth growth.


\section{III. Basic Summary of the Data}

Figure 2 in the Appendix gives a table of basic descriptive statistics for the ToothGrowth dataset. The mean guinea pig had tooth length of 18.81 mm (I assume this is mm. The help documentation does not say.). There were an equal numbers of 0.5, 1.0 and 2.0 mg doses administered. The standard deviation of tooth length is 7.65 mm.

The tooth length data do not appear to be normally distributed when looking at the histogram in Figugre 3. However, a Jarque-Bera test of tooth length suggests that the distribution of tooth length does roughly conform to normality. 

\section{IV. Using Hypothesis Tests and Confidence Intervals}

In an attempt to compare the means of tooth length along two dimensions (supplement type and dosage), I created the following table of means and 95% confidence intervals.


```{r echo=FALSE}
bySupp_OJ <- ToothGrowth[ToothGrowth$supp=="OJ",1]
bySupp_VC <- ToothGrowth[ToothGrowth$supp=="VC",1]
byDose_05 <- ToothGrowth[ToothGrowth$dose==0.5,1]
byDose_10 <- ToothGrowth[ToothGrowth$dose==1.0,1]
byDose_20 <- ToothGrowth[ToothGrowth$dose==2.0,1]
meantests <- data.frame(0,nrow=5, ncol=3)
meantests[1,1] <- mean(bySupp_OJ); meantests[1,2] <- mean(bySupp_OJ)-qnorm(0.975)*sd(bySupp_OJ)/sqrt(length(bySupp_VC)); meantests[1,3] <- mean(bySupp_OJ)+ qnorm(0.975)*sd(bySupp_OJ)/sqrt(length(bySupp_OJ)) 
meantests[2,1] <- mean(bySupp_VC); meantests[2,2] <- mean(bySupp_VC)-qnorm(0.975)*sd(bySupp_VC)/sqrt(length(bySupp_VC)); meantests[2,3] <- mean(bySupp_VC)+ qnorm(0.975)*sd(bySupp_VC)/sqrt(length(bySupp_VC)) 
meantests[3,1] <- mean(byDose_05); meantests[3,2] <- mean(byDose_05)-qnorm(0.975)*sd(byDose_05)/sqrt(length(byDose_05)); meantests[3,3] <- mean(byDose_05)+ qnorm(0.975)*sd(byDose_05)/sqrt(length(byDose_05)) 
meantests[4,1] <- mean(byDose_10); meantests[4,2] <- mean(byDose_10)-qnorm(0.975)*sd(byDose_10)/sqrt(length(byDose_10)); meantests[4,3] <- mean(byDose_10)+ qnorm(0.975)*sd(byDose_10)/sqrt(length(byDose_10)) 
meantests[5,1] <- mean(byDose_20); meantests[5,2] <- mean(byDose_20)-qnorm(0.975)*sd(byDose_20)/sqrt(length(byDose_20)); meantests[5,3] <- mean(byDose_20)+ qnorm(0.975)*sd(byDose_20)/sqrt(length(byDose_20)) 
row.names(meantests) <- c("Supp = OJ", "Supp = VC", "Dose = 0.5 mg", "Dose = 1.0 mg", "Dose = 2.0 mg")
names(meantests) <- c("Mean", "Lower Limit (95%)", "Upper Limit (95%)")
meantests
```

There is very little overlap in the 95% confidence interval for the means of the OJ vs. VC supplement groups, suggesting that their means are statistically different. Similarly, as can also be observed in Figure 1, the means of tooth length for the three different dosages are also statistically different from each other.

\section{V. Necessary Assumptions}

The confidence intervals above were constructed using the 95% two-sided (97.5% one-sided) quantile of the normal distribution. Therefore, the underlying assumption of this analysis is that the data are drawn from a normal distribution centered at the respective sample means and with variance approximated by the sample variances.

In order to interpret measures of differences in tooth *length* as representative or changes in tooth *growth*, it was also necessary to assume that pre-treatment tooth length for all subjects was roughly equal or that, at the very least, subjects were randomly placed into various treatment groups. 

\section{VI. Conclusion}

This brief and informal analysis gives some preliminary evidence that the Vitamin C treatment did have a statistically significant impact on tooth growth. Subjects who received higher doses experienced greater tooth lengths.

The R markdown document with the underlying code used to generate this report can be found at [this GitHub repo](https://github.com/JayLamb20/MOOC_Stuff/tree/master/Statistical_Inference).

\pagebreak


\appendix
\section{Appendix: Plots and Result Tables}

\subsection{Figure 1. Boxplot of Tooth Length by Dose}

```{r messages=FALSE}
library(datasets)
data(ToothGrowth)
```

```{r echo=FALSE}
par(mfcol=c(1,1))
par(mar=c(5,5,5,5))
ToothGrowth$dosefac <- as.factor(ToothGrowth$dose)
boxplot(formula = ToothGrowth$len ~ ToothGrowth$dosefac, ylab="Tooth Length", xlab="Dose", main="Boxplot of Tooth Length by Dose")
```

\subsection{Figure 2. ToothGrowth Descriptive Statistics}

```{r echo=FALSE}
library(datasets)
data(ToothGrowth)
library(pastecs)
```

```{r echo=FALSE}
stat.desc(ToothGrowth)
```

\subsection{Figure 3. Histogram of Tooth Length}

```{r echo=FALSE, message=FALSE}
library(tseries)
library(timeDate)
```

```{r echo=FALSE}
hist(ToothGrowth$len, main="Tooth Length", col="green")
noquote(paste("Jarque-Bera Test |","JB Stat", round(jarque.bera.test(ToothGrowth$len)$statistic,2),"p-value",round(jarque.bera.test(ToothGrowth$len)$p.value,5), sep= " "))
noquote(paste("Excess Kurtosis =", round(kurtosis(ToothGrowth$len)[1],5),"| Skewness =",round(skewness(ToothGrowth$len)[1],5), sep= " "))
```

\pagebreak
\begin{thebibliography}{9}    

\bibitem{pastecs}
    Phillippe Crosjean et. al,
    "Pastecs: Package for Analysis of Space-Time Ecological Series",
    \emph{CRAN},
    2014.
    Retrieved from: http://bit.ly/1pnSI2r
    
\end{thebibliography}