<center> <h2>Interactive Python - Week 4</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 5.1 - Introduction*

**A. Expectations**

- you should know how the class works
- you should know some Python!
- part one will stay open for the duration of this class
- we're going to add more complexity this time
- the final project = "Rice Rocks"

####*Video 5.2 - Mouse Input*

**A. Gaming**

- three excuses for losing at a video game:
    1. I lagged
    2. Mom distracted me!
    3. I got hacked! Yeah!

**B. Handler for clicks**

- the mouseclick handler takes one parameter(position)
    - the position is a tuple of two integers
    - remember that tuples are immutable

####*Video 5.3 - List Methods*

**A. List methods**

- you can use the ```in``` operator to check if stuff is in there
    - for example ```if 4 in mylist:```
- you could use ```index()``` to find where stuff is in the list
    - for exmplae ```position = mylist.index("apple")```
- Using ```pop``` is like "cut"
    - returns the element you poppeds
- ```append``` adds objects, ```extend``` adds elements
    - [see StackOverflow for more](http://stackoverflow.com/questions/252703/python-append-vs-extend)
- lists are really powerful!
    - they are "the workhorse of Python"
    
####*Video 5.4 - List Examples*

**A. More Examples**

- if you're iterating over a list, you cannot remove items from the list inside that loop
    - you need to create a temp list of things to remove and remove them at the end of the for loop

####*Video 5.5 - Iteration*

**A. Stuff**

- it might be better to use ```remove()``` in a list rather than indices (popping with indices will blow up your indexing scheme!)

####*Video 5.6 - Dictionaries*

**A. Intro to Dictionaries**

- a dictionary is a mapping consisting of key-value pairs
- in a list, indices are fixed and numbered
    - in a dictionary, you can use anything as keys (even non-numeric stuff)
    - could even map a key to a list
- when you iterate over a dictionary, you get all the keys
    - no guarantees about order, but you will get all of them once
- you can loop over dictionary items:

```{r eval=FALSE}
for key, value in MYDICTIONARY.items():
```

**B. Random Ordering**

- you can use ```random.shuffle()``` to randomly rearrange a list

####*Video 5.7 - Loading Images*

**A. Using Images**

- you can load an image with ```im = simplegui.load_image(URL)```
- remember that loading images takes time
    - the image will be cached once you've downloaded it once
- ```canvas.draw_image(im, src_center, src_size, dst_center, dst_size)```
- the last 4 parameters will all be tuples
    - CodeSkupltor won't stop running if image load fails, but know that failure can happen
- Map magnifier:

```{r eval=FALSE}
# First draws entire map at reduced resolution
# Second draws magnifier pane around clicked area


# Demonstration of a magnifier on a map

import simplegui

# 1521x1818 pixel map of native American language
# source - Gutenberg project

image = simplegui.load_image("http://commondatastorage.googleapis.com/codeskulptor-assets/gutenberg.jpg")

# Image dimensions
MAP_WIDTH = 1521
MAP_HEIGHT = 1818

# Scaling factor
SCALE = 3

# Canvas size
CAN_WIDTH = MAP_WIDTH // SCALE
CAN_HEIGHT = MAP_HEIGHT // SCALE

# Size of magnifier pane and initial center
MAG_SIZE = 120
mag_pos = [CAN_WIDTH // 2, CAN_HEIGHT // 2]


# Event handlers
# Move magnifier to clicked position
def click(pos):
    global mag_pos
    mag_pos = list(pos)

# Draw map and magnified region
def draw(canvas):
    # Draw map
    canvas.draw_image(image, 
            [MAP_WIDTH // 2, MAP_HEIGHT // 2], [MAP_WIDTH, MAP_HEIGHT], 
            [CAN_WIDTH // 2, CAN_HEIGHT // 2], [CAN_WIDTH, CAN_HEIGHT])

    # Draw magnifier    
    map_center = [SCALE * mag_pos[0], SCALE * mag_pos[1]]
    map_rectangle = [MAG_SIZE, MAG_SIZE]
    mag_center = mag_pos
    mag_rectangle = [MAG_SIZE, MAG_SIZE]
    canvas.draw_image(image, map_center, map_rectangle, mag_center, mag_rectangle)
    
# Create frame for scaled map
frame = simplegui.create_frame("Map magnifier", CAN_WIDTH, CAN_HEIGHT)

# register even handlers
frame.set_mouseclick_handler(click)    
frame.set_draw_handler(draw)

# Start frame
frame.start()
```

- think about scaling up to "undo" the downscaling of the canvas

####*Video 5.8 - Visualizing Iteration*

**A. Notes**

- list comprehension...how the cool kids do it:

```{r eval=FALSE}

## Squaring with For loop
def square_list1(numbers):
    """Returns a list of the squares of the numbers in the input."""
    result = []
    for n in numbers:
        result.append(n ** 2)
    return result
    
## Squaring with list comprehension
def square_list2(numbers):
    """Returns a list of the squares of the numbers in the input."""
    return [n ** 2 for n in numbers]

def is_in_range(ball):
    """Returns whether the ball is in the desired range.  """
    return ball[0] >= 0 and ball[0] <= 100 and ball[1] >= 0 and ball[1] <= 100

# squaring with loop
def balls_in_range1(balls):
    """Returns a list of those input balls that are within the desired range."""
    result = []
    for ball in balls:
        if is_in_range(ball):
            result.append(ball)
    return result

# squaring with list comprehension
def balls_in_range2(balls):
    return [ball for ball in balls if is_in_range(ball)]

```

- list comprehension allows you to "map" elements (do stuff to them) and to "filter" (return a list of only some subset of the list you loop through)
- you can use the ```+=``` operator with strings!!  (makes sense, still cool)

####*Video 5.8 - Programming Tips**

**A. Errors with Lists**

- remember that indexing starts at 0!
    - ```mylist[len(mylist)]``` is not going to work for you

**B. Dictionaries vs Lists**

- dictionaries have key-value relationships
- lists are just collections of elements
- indices of a list must be contiguous integers
- dictionary keys can be anything
- you can slice lists but not dictionaries
    - there's no ordering of the dictionary, so slicing is useless
- lists and dictionaries can be mutated
- what can be a key?
    - boolean, string, tuple, dictionary...NOT a list
    - has to be something hashable
    
**C. Loading Images**

- note that CodeSkulptor will not error with a bad URL
    - programs will return 0 if you reference height and width when the image hasn't loaded yet

####*Video 5.9 - Memory*

**A. How it works**

- if cards match, leave them face up
- if they don't, cover them up
- this is a difficult project
- NOTE: ignore clicks on cards that were already exposed
- consider using a global variable for "how many cards are flipped"
    - handle checking if values are the same inside the "state = 2" bit of buttonclick()...the mouse handler
-