<center> <h2>Reproducible Research - Week 2</h2> </center>
<center> <h3>---------------------------------------</h3> </center>

####*Video 2.1 Coding Standards in R*

**A. Why Should We Care?**

- standards make your code more readable
- this is just about some basic, minimal standards for R
1. Principle 1: Write Your Code In a Text Editor, Saved as a Text File (maybe txt, ascii)
2. Principle 2: Indent Your Code
    - in RStudio, change prefs by going to File --> Preferences --> Code Editing
    - maybe 8 space indent plus character width
    - this might stop you from doing weird things like quadruple-nested for loops
    - this prevents you from making fundamental mistakes (writing unreadable things)
3. Principle 3: Limit the Width of your code (perhaps to 80 columns)
4. Principal 4: Limit the length of individual functions
    - limit your functions to a single, basic activity
    - break separate things (read data, fit a model, whatever) into separate pieces
    - you should be able to see the code for your function on one screen of the editor
    - makes debugging/traceback easier
    - debugging typically stops at telling you "there's a problem somewhere in this function"
    
####*Video 2.2 Markdown*

**A. Why Use Markdown: Basics**

- it's a simplified markup language.  You can focus more on content than on complicated tags
- all the course slides are written in R Markdown

**B. Introduction to Markdown**

- "a text-to-HTM conversion tool for web writers"
- Italics --> single asterisk, e.g. *this is italics*
- Bold --> Double asterisk, e.g. **this is bold**
- headings, uses "#". More equals deeper-level heading
- ordered lists with numbers, unordered lists with dashes
- Your number in ordered lists actually don't matter...you just need to use A number
- links? 
```{r eval=FALSE, echo=TRUE}
[text](url)
```

- a better way to do links:

```{r eval=FALSE, echo=TRUE}
    - "I spend so much time reading [R bloggers][1] and [Simply Statistics][2]"
    - then, later on:
    - "[1] URL 1"
    - "[2] URL 2"
```

- New lines:
    - new lines require a double space after the end of a line
- other links:
    - [the official Markdown documentation](http://daringfireball.net/projects/markdown/syntax)
    - [Github Markdown guide](https://help.github.com/articles/github-flavored-markdown/)
    
####*Video 2.3 R Markdown Lecture*

- RStudio easily handles the R markdown --> markdown --> html work flow
- "these slides were written in R and converted to slides using the slidify package"
- you never edit the markdown file or HTML...just the R markdown file


####*Video 2.4 R markdown Demonstration*

- this lecture was borrow from the R Programming course

####*Video 2.5 knitr (pt 1)

**A. Problems, Problems**

- authors have to put in a lot of effort to put data/results on the web
- readers have to download data/results individually and piece together what happened
- solution: put data, code, text all in a single document

**B. Literate Statistical Programming**

- comes to us from Stanford's Don Knuth
- have code and analysis there togethers
- article = stream of text and code
- "literate programs are **weaved** into human-readable, **tangled** into computer readable"

**C. How do I make My Work Reproducible?**

- decide to do it!
- use version control to tracks changes as you go along
    - GIT, SVN
- Use statistical software with code
    - pick-and-click through GUI not reproducible
- don't save any output (intermediate output)
    - it might be better to just give raw data and pre-processing code
- save data in shareable, non-proprietary formats

**D. Benefits of Literate Programming**

- text and code all in one place (logical order)
- data and results are automatically updated to reflect changes
    - the whole document will automatically update itself
- code is live -- automatic "regression test" when building a document
    - if you put in mistakes, you know you put in errors
    - doing this a lot keeps yoru errors from piling up
    
**E. Literate Programming: Cons**

- text and code all in one place, so the code can obscure human readable
- If your analysis is really complicated, this whole compiling thing can slow you down a lot

####*Video 2.6 knitr (pt 2)*

**A. What is knitr?**

- An R package, built into RStudio GUI
- supports RMarkdown, LaTeX, and HTML
- can export PDF, HTML

**B. Markdown vs. Markup**

- HTML, LaTeX = a "markup" language using tags
- markdown is way simpler, mostly just read text
- simple, intuitive formatting elements
- no need for a special editor

**C. What is knitr Good For?**

- manuals
- tutorials
- reports (esp. if generated periodically)
- data preprocessing documents/summaries

####*Video 2.7 knitr (pt 3)*

**A. Making Your First Document**

- some notes in this lecture's slides for how to replicate knitr outside of the RStudio environment
    - use knit2html() and browseURL()
- code chunks can have names, which can be useful when you start making graphics. think --> {r chunkname}
- code in a code chunk is echoed by default, as are calculations

####*Video 2.8 knitr (pt 4)*

**A. Processing of knitr Documents**

- you write the RMarkdown (.RMD)
- knitr produces a Markdown document (.md)
- knitr converts the markdown document into HTML (by default)
    - .RMD -> .md -> HTML
    
**B. An Example**

- to prevent it from showing the code (just show result), use echo=FALSE in the curly brackets
- hide the results with results="hide" or eval=FALSE options
- you can write numbers into your sentences!!!!

```{r computetime, echo=FALSE}
time <- format(Sys.time(), "%a %b %d %X %Y")
rand <- rnorm(1)
```

- The current time is `r time`. My favorite random number is `r rand`.

**C. Incorporating Graphics**

- the HTML file stands alone. It doesn't rely on other image files
    - the full image is embedded, not calling on some file stored elsewhere
    - not that efficient, but nice for transferability

**D. Making Tabls with xtable**

```{r loadpackages, echo=FALSE, message=FALSE, warning=FALSE}
#load packages in a separate chunk, so we can avoid printing package-loading messages
library(datasets)
library(xtable)
```

```{r fitmodel, results="asis"}

#using the option results="asis" in the curely brackets interprets xt output in a pretty way (see below)

#Get some data
data(airquality)

#estimate a regression
fit <- lm(Ozone ~ Wind + Temp + Solar.R, data = airquality)

#make the output pretty
xt <- xtable(summary(fit))
print(xt, type = "html")
```

- the xtable package lets you make some nicer tables out of stuff

**E. Setting Global Options for code chunks**

- sometimes we want to set global options
- you have to write some code for global options
- look up how to do this
- setting global options can save you a bit of time and effort
- "asis" means just show the raw results
- you can modify figures with fig.height, fig.width

**F. Caching Computations**

- what if one chunk takes a long time to run?
- all chunks have to be re-computed every time you re-knit the file
- the cache=TRUE option can be set on a chunk-by-chunk basis to store resuls of computation
- after the first run, results are loaded from cache (way faster than re-doing computations)
- caching won't explicitly check inter-chunk dependencies
    - so make sure you re-estimate everything affected by a change
    - chunks with significant side effects (effects outside document) may not be cacheable

####*Video 2.9 Introduction to Peer Assessment 1*

**A. The Assignment**

- do two things:
    - do a bit of data analysis
    - write up a report using RMarkdown and knitr
- to submit:
    - check in to a GitHub repository
    - push it there
- submission stuff:
    - GitHub Repo URL
    - SHA-1 Commit Hash (so we know which version to look at)

**B. Finding the GitHub Repository URL**

- The repo is just up at the top
    - github.com/name/reponame
    
**C. Finding the SHA-1 commit hash**

- click on where it says "## commits"
- It will have tones of them, with the most recent at the top
- on the very right, there's a shortened version of the SHA-1 hash
    - click the little clipboard symbol, then you can paste right into Coursera

**D. Browsing a Specific Commit**

- go to the repo URL on GitHub
- go to the commits
- click "browse code"
    - this will show you the files at that snapshot
- 